# Introduction {#intro}
The evolutionary goal of a species is to survive. This single goal is achieved by adopting to the environment and by optimising strategies to live in the specific environment and reproduce [@darwin_origin_1859; @raup_biological_1986; @raup_role_1994]. Higher organisms therefore developed structures which are able to help with optimising their behaviour on different levels. Very early on developed concepts were mechanosensation (touch, hearing), olfaction based on chemotaxis, and vision [@schlosser_short_2018]. 

Information which is accumulated over the course of time including sensory stimuli have to be stored in different areas. The coordination of areas and timing of cells can be crucial for memory formation and can be achieved by oscillatory activity ###Bressler93varela2001. Oscillations arise in the brain through changes induced by population activity, synaptic inputs, but also dendritic depolarisation and can be measured as local field potentials (LFP) ##herrera2016. 

One important higher level function, which is not solely a processing of sensory inputs, is the computation for spatial navigation. Navigation can be observed in several species of the animal kingdom such as ants [@wehner_desert_2003], drosophila (cite), pigeons, rodents, bats, and dolphins [@dacke_honeybee_2007] and is essential to find food, a partner, or avoiding danger. 


Over the last decades the endeavour to understand the *"cognitive map"* [@tolman_cognitive_1948] and how we use the abstraction of space (citation needed) drove research and revealed several concepts of how space can be deconstructed (place, speed, grid, head-direction, vectors, border etc) and represented by the brain. Coding space properties can be achieved by different mechanisms allowing for a detailed representation of the environment. If we take place cells, they can be found to fire with a higher rate in specific position which are referred to as place fields ###dostrovsky71, mcnaughton76/78. This principle of coding properties with the count of spikes is called rate-code.

However, this is not the only possible way of transferring information. An additional way a spike can code space is via the phase-code ###skaggs96, mcnaugthon, mehta. The phase-code is based on the fact that inputs and outputs can shape the electric field in the brain. The resulting field potential can show oscillatory behaviour, meaning a sine-wave at specific frequencies which relate to the rhythmic generation of the signal. If one now aligns the occurrence of spikes to the phase, the position on a sine-wave, it can be observed that location can also be reflected by the spike phase and therefore can code additionally information via the same spike. This was found to play an important role in several areas of the brain (LS, MEC, CA1 ####). On of the major frequency bands involved in the phase code is the $\theta$ band.


On a technical level new tools are available to understand the brain a bit better. This also includes ways of analysing data and to estimate properties of different measurements. In the past, until the beginning of the 20th century, Bayes' theorem was used at large to estimate parameters in astrology ###Gauss and other notions of events. However, the use of Bayes' theorem came out of fashion with Fisher ####1922 establishing the nowadays know frequentist hypothesis testing due to the fact that more complex problems could be solved easier using maximum likelihood estimation. This changed over the last decades with the implementation of Markov chain Monte Carlo simulations on modern computers. More complex issues can again be framed in Bayes' theorem without solving manually for complex derivatives which can become more difficult in higher parameter space and might be impossible to do by hand ####Carpenter2015 "Stan. The interpretation of estimates derived using Bayes' give a complete distribution for the estimates allowing to quantify uncertainty. 

## Theta
$\theta$ oscillations range from 4 to 12 Hz ###buzsaki2002 and were for the first time observed in the hippocampus and septum of rats and cats by ###jung&Kornmüller1938 and later by @green_hippocampal_1954 in rabbits, and monkeys. Exposure of different stimuli and lead to different oscillatory responses including slow oscillations in the $\theta$ range in the hippocampal area but also in the medial septum. 

A decade later $\theta$ was reported in the context of navigation for the first time [@vanderwolf_hippocampal_1969]. Freely moving rats show synchronised hippocampal activity reflected in the LFP during exploration [@winson_patterns_1974]. Similarly these oscillations can be found in the mouse and also human, however, in humans with lower frequency [@ekstrom_human_2005; @miller_lateralized_2018; @watrous_comparative_2013]. Interestingly enough, bats do not show a classical $\theta$ pattern during exploration ##eliav2018. Further, $\theta$ activity in rodents is almost constantly occurring during movement compared to burst activity in humans [@ekstrom_human_2005; @watrous_comparative_2013]. 

Additionally to exploration and movements, the presence of $\theta$ is crucial for working memory where it is involved in synchronising cells #Lee2005 jensens2002, the coordination of activity across different areas ###liebe2004#sarnthein1997  but also for spatial memory #wiston78.

It is also important to consider that $\theta$ oscillations in humans can be divided in two different groups (high: 2-4Hz; low: 4--14Hz) which occur during different conditions ###Abhinav Goyal. High $\theta$ would be more likely to correlate with movement and was suggested to be more relevant to spatial processing in humans.
Mechanistically $\theta$ oscillations can be differentiated into two different types one of which is atropine sensitive, slightly slower ($4-7Hz$), and associated with rest or exists under anaesthesia. A non-sensitive $\theta$ which is slightly faster ($7-12Hz$) occurs during running and exploration ###kramis75.

Cortical structures are known *"current generators"* but not yet shown to be able to generate theta by themselves *in-vivo*. Origin of theta is believed to be in deeper regions such as the septum ### petsch1962
Origin of $\theta$ is not completely resolved but some areas have the potential to generate $\theta$ locally ###Goutagny2009,chapman2007, however, evidence for local generation *in-vivo* is sparse. Evidence from *in-vitro* studies suggests that connectivity between glutamergic cells and inhibitory interneurons can be sufficient to generate $\theta$ in the hippocampus. 




HCN channels causing $I_{h}$ could play an important role in oscillatory behaviour and are expressed in the medial septal neurons with firing activity in the $\theta$ band ###Hangya2009Varga2008.

$\theta$ cycles can be used to code for distinct stimuli from the environment.

One implication of the oscillation is that some cells fire according to the phase of the oscillations. In the hippocampus it can be observed that pyramidal cells ride on the $\theta$ oscillation which results in modulation of spike timing ### mehta okeefe93. 

Further, it was proposed by Lisman1995 that information can be nested in $\theta$ cycles via $\gamma$-oscillations


Already early on there was a consensus that oscillations are associated with navigation in behaving rodents [@okeefe_hippocampus_1971; @vanderwolf_hippocampal_1969]

(ref:quoteKD) --- Footnote by J.O'Keefe & J.Dostrovsky , 1971

(ref:wholeQuoteKD) * Vanderwolf has reported (and we agree) that the former behaviours are associated with $\theta$ activity in the hippocampal EEG, while, during the latter behaviours, the EEG shows irregular slow waves.

> (ref:wholeQuoteKD)
>
> `r tufte::quote_footer('(ref:quoteKD)')`



Today 40 years later, however, the exact purpose of $\theta$ is still not completely resolved. 



## Medial Septum
The medial septum (MS) a structure located in the forebrain (Figure \@ref(fig:MS-Scheme)) is assumed to be a central $\theta$ generator ###petsche62. Due to connectivity and oscillatory properties of medial septal cells it is crucial to entrain $\theta$ in different areas of the brain synchronising activity #####gray79 bland79, petsche62. Interfering with this important synchronisation will have decisive consequences for learning and memory ###mizumor 79, chrobak89 ##winston78. *In-vivo* the medial septum is a crucial part of $\theta$ generation and precedes hippocampal $\theta$ onset by roughly $500ms$ ###bland1999 indicating that it is to some degree primarily driving synchronicity.

Deactivation of the MS will lead to spatial deficits



(ref:MS-SchemeShortCaption) Schematic of medial septum
(ref:MS-SchemeCaption) **Schematic of medial septum.** Sagittal (A) and coronal (B) section. Blue shows the location of the MS, light gray represents ventricular space and dark gray fibre bundels. The dashed line in (A) indicates the position of the section of (B). Modified from Allen Brain Atlas.
```{r MS-Scheme, fig.cap = '(ref:MS-SchemeCaption)', fig.scap='(ref:MS-SchemeShortCaption)', fig.pos='H', fig.align='center', echo=FALSE, fig.height=3, fig.asp=0.4}
sagMS  <- grImport2::pictureGrob(grImport2::readPicture(rawToChar(rsvg::rsvg_svg("Figures/Schemes/MS_sagittal_clean.svg"))))
coronalMS  <- grImport2::pictureGrob(grImport2::readPicture(rawToChar(rsvg::rsvg_svg("Figures/Schemes/MS_coronal_clean.svg"))))
wrap_elements(gridExtra::arrangeGrob(sagMS), clip = T) + wrap_elements(gridExtra::arrangeGrob(coronalMS), clip=T) + plot_layout(widths = c(3.2,2)) &
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size=24))
```

The MS is a region which projects to several areas of the brain including significant projections to the hippocampus [@fuhrmann_locomotion_2015; @huh_glutamatergic_2010; @joshi_behavior-dependent_2017; @robinson_optogenetic_2016; @unal_spatio-temporal_2018; @unal_synaptic_2015; @vandecasteele_optogenetic_2014] and the parahippocampal network [@alonso_study_1984; @desikan_target_2018; @fuchs_local_2016; @gonzalez-sulser_gabaergic_2014; @van_groen_connections_1990] where it targets preferred cell types [@desikan_target_2018; @fuchs_local_2016; @fuhrmann_locomotion_2015; @gonzalez-sulser_gabaergic_2014; @unal_synaptic_2015] and 

In this study I will focus on GABAergic and cholinergic projections.

### Gabaergic projections
The majority of the medial septum, however, is made up of GABAergic cells ##freund1989PV which in return can be divided into several sub-classes ### somogyi. Generally these interneurons are thought to fire rhythmically in theta bursts. Two major groups were identified by @joshi_behavior-dependent_2017. One of the two was referred to as "Teevra" cells, a PV^+^ cluster firing in strong bursts and targeting PV^+^ and CCK^+^ cells in the CA3. However, Teevra cells were not found to project to the parahippocampal network or the retrosplenial cortex suggesting that they play a defined role in modulation of the hippocampus @joshi_behavior-dependent_2017. "Komal" cells, a second group of PV^+^ cells in the medial septum, don't seem to be modulated by running speed. An additional difference between these two cell types is the CA1 theta phase in which the spikes occur. Komal cells fire preferentially at the peak of the theta phase whereas Teevra cells are locked at the trough. @varga_presence_2008 showed that PV^+^ cells expressing HCN channels are more likely to fire at trough of hippocampal $\theta$. Expression of HCN also lead to a more bursty behaviour of cells suggesting that HCN expression corresponds to Teevra cells. In line with these observations by @xu_hippocampal_2004 where they showed a reduction of hippocampal $\theta$ after injecting an HCN blocker into the MS. Recently it was shown by @salib_gabaergic_2019-1 that also GABAergic low-rhythmic-firing neurons (LRN) in the medial septum do project to cortical areas and get inhibitory feedback during hippocampal sharp-waves. @salib_gabaergic_2019-1 postulates that PV^+^ LRNs might be more involved in medial septal local connectivity opposed to septal PV^-^ LRNs which target the DG and CA3.
Targets of PV^+^ cells in the hippocampus or MEC are mainly interneurons which partially express PV themselves [@fuchs_local_2016; @gonzalez-sulser_gabaergic_2014; @unal_synaptic_2015]. Interestingly enough no layer 1 interneurons are reported to receive inputs in cortical areas from PV^+^ septal projections. Considering that Teevra cells are modulated by speed it is important to note that a stimulation at $\theta$ frequency of PV^+^ septal fibres projecting to the hippocampus can reduce running speed ###Bender2015. Further, it did not change spatial firing patterns, nor did it change the firing frequency of hippocampal cells.

### Cholinergic projections
Cholinergic cells (ChAT^+^) in the MS are thought be slowly firing  [@manseau_hippocamposeptal_2008-1] and project to the hippocampus and MEC [@desikan_target_2018; @unal_synaptic_2015], Similarly to the PV^+^ cells in the MS ChAT^+^ cells project mainly to distinct areas. For example, only a few cells will project at the same time to the CA1 and the MEC [@unal_synaptic_2015]. The majority of cells, however, will target specifically one of the two areas. Cholinergic projections can work in different ways due to the properties of acetylcholine which can slowly inhibit, slowly excite by binding muscarinic receptors or it can excite a cell in a fast way via nicotinic receptor binding ### citations. All these mechanisms were shown to be used by cholinergic projections from the MS. In the entorhinal cortex the majority of cells receiving nicotinic inputs were superficial layer 1 interneurons and layer 2 p5HT~3~R^+^ cells [@desikan_target_2018]. Hyperpolarisation due to probable activation of muscarinic receptor (Type II) were observed in pyramidal cells, stellate cells and their intermediates as well as putative PV^+^ interneurons in the MEC. A similar pattern was observed in the lateral entorhinal cortex (LEC). It was also reported that projections to DG can inhibit granule cells by activating astrocytes which in turn will excite hilar interneurons and mossy cells through glutamate  [@pabst_astrocyte_2016]. When interfering with cholinergic septal cells using a chemogenetic approach to increase cell firing $\theta$-power was reduced in the MEC, however, grid cell firing in terms of stability or grid score was not affected [@carpenter_modulating_2017]. In comparison activation of cholinergic septal cells lead to to reduced $\theta$-power and a slight increase in coherence in the hippocampus both of which also seem to be state dependent [@vandecasteele_optogenetic_2014]. It is important to note though that an activation of cholinergic cells does not affect the field potential in the MS, but will increase speed of running rats and increase correlation between $\theta$ and running speed.

<!-- ### Glutamergic projections  -->
<!-- Glutamergic projections were found in the hippocampus and target -->
<!-- @robinson_optogenetic_2016 showed *in-vitro* that glutamergic cells in the MS target local MS GABAergic interneurons but also some cholinergic interneurons. Measuring hippocampal cells revealed that only a small fraction of cells were functionally connected and which was in line with a very spars projection pattern. @robinson_optogenetic_2016 showed further that upon activation of these fibres *in-vivo* hippocampal $\theta$ could only be induced when stimulated in the MS or stimulating non-glutamergic fibres in the hippocampus. Similarly, @gonzalez-sulser_gabaergic_2014 reported very low connectivity rates for glutamergic responses in the MEC. These studies suggest a more local importance for glutamergic MS cells. In contrast @fuhrmann_locomotion_2015 reported functional connectivity to hippocampal interneurons with an increased connectivity rate compared to @robinson_optogenetic_2016 (4% vs 28%). Additionally they record *in-vivo* hippocampal LFP and stimulate glutamergic cells in the MS using channelrhodopsin. They report that after blocking local glutamergic connectivity in the MS by injecting the glutamate blocker NBQX into the septum optogenetic frequency responses could be strongly reduced, however, it also reduced the correlation between running speed and $\theta$ suggesting a role of glutamergic projections. Additionally, they find that upon light stimulation in the MS mice started to run during the experiment suggesting a role in locomotion similarly to what was reported for cholinergic septal cells [@vandecasteele_optogenetic_2014]. Similar findings by ###Justus in the MEC indicate that glutamergic projections to the MEC are also likelier than expected from previous data [@gonzalez-sulser_gabaergic_2014]. Not only do glutamergic cells project to MEC but imaging of axon terminals reveals a strong positive correlation between speed and fibre activity. However, glutamergic cells in the MS are not uniquely positively correlated with speed but represent a heterogeneous population ###Justus.  -->



Functionally, the medial septum is highly implicated in being a "theta generator"

Furthermore, an inactivation of the MS in rats using lidocaine or muscimol will not only lead to reduced $\theta$ activity but also to reduced grid cell stability in the MEC and destabilise place fields in the hippocampus [@brandon_new_2014; @koenig_spatial_2011-1]. Head-direction cells on the other hand were not significantly effected. Recently, @lepperod_optogenetic_2021 showed in rats that activation of PV^+^ cells in the MS could drive LFP in the MEC, reduced $\theta$ and phase precession in grid cells but did not abolish grid cell stability. Furthermore, it was shown that speed modulation which normally occurs in the MEC [@gonzalez-sulser_grid_2017] was only disrupted on the cell LFP level due to MS stimulation, but not on the cell level.

It could be shown by @vann_using_2000:1 using FOS-activity as a molecular marker that lesioning the fornix, the output fibre track of the MS, will strongly reduce cell activity in the PaS, hippocampus, and other parahippocampal areas during a radial arm maze task. Another lesion study by @jacob_medial_2017:1 revealed that lesion of the MS will impair spatial learning. To test this rats were trained to run a linear track and stop at a visually indicated distance to receive a reward. After removing the visual cue rats were able to estimate the distance correctly, however, showed an increased error in estimation when the MS or the MEC were lesioned. A lesion in the hippocampus on the other hand showed no significant loss in accuracy. Findings supporting the importance of the MS for distance estimation were found in humans where $\theta$-bursts could be observed before navigating to a target location .The observed bush et al 

Therefore, a major focus of research was to investigate projections to the hippocampus or MEC. However, early evidence indicated that also other regions of the parahippocampal area are innervated by the medial septum [@swanson_connections_1979]. Recently this changed and an increased interest into the parasubiculum (PaS) gained traction giving the first evidence of MS GABAergic projections to the PaS [@tang_functional_2016; @unal_synaptic_2015].


## Parasubiculum
The parasubiculum on the other hand is part of the parahippocampal formation (Figure \@ref(fig:PaS-Scheme)), the posterior part of the mouse brain [@boccara_threeplane_2015], and is found across different species [@ding_comparative_2013]. The relatively small size and orientation of the PaS rendered recordings in-vivo challenging but were made possible recently
[@boccara_grid_2010; @kornienko_non-rhythmic_2018; @tang_functional_2016; @taube_place_1995] ##ebbesen. It is adjacent to the medial entorhinal cortex and the presubiculum which, as the PaS, have spatially linked functional cell types coding for different aspects used for navigation [@boccara_grid_2010] . The PaS has a high density of Wolframin ER Transmembrane Glycoprotein (WFS1) positive cells which can also be found in pyramidal cells of the MEC layer 2 and pyramids of the CA1 [@kitamura_island_2014; @luuk_distribution_2008; @ramsden_laminar_2015; @ray_grid-layout_2014; @sammons_electrophysiological_2019] . They represent the largest proportion of cells in the PaS and are until now the major excitatory cell type  [@sammons_electrophysiological_2019]. The close vicinity to the MEC make it also a good canditated relaying information from the hippocampus to the MEC. Indeed, projections from the PaS were found to target several layers of the MEC mainly exciting their postsynaptic cells [@canto_all_2012]. Inhibition could only be detected in a small fraction of cells underlining the importance of pyramidal cells in the PaS. Recently, it was shown that local parasubicular interneurons are likely to project to WFS1 cells (fast spiking interneurons: 44%; not fast spiking interneurons: 13%) whilst receiving glutamergic inputs from WFS1 cells with a high probability (fast spiking interneurons: 23%; not fast spiking interneurons: 2.6%) [@sammons_local_nodate]. This feedback circuit sets up the PaS as a oscillatory network. Indeed the PaS was shown to have strong $\theta$-oscillations *in-vivo* [@boccara_grid_2010; @tang_functional_2016]. Additionally to the circuitry WFS1 cells have an intrinsic oscillatory behaviour and can be further modulated by acetylcholine [@glasgow_muscarinic_2013; @salkoff_synaptic_2015; @sammons_electrophysiological_2019].

Due to its close vicinity to the MEC 

(ref:PaS-SchemeShortCaption) Schematic of parasubiculum
(ref:PaS-SchemeCaption) **Schematic of parasubiculum.** A sagittal (A) and horizontal (B) section. Turquoise shows the location of the PaS, light grey represents ventricular space and dark grey fibre bundles. The dark green in (B) represents the pyramidal cell layer II of the medial entorhinal cortex. The cornu ammonis and the dentate gyrus are marked in light blue. (C) A stained horizontal section showing neurons labelled with NeuN (top) and WFS^+^ pyramidal cells in the PaS, MECII, and CA1 of the hippocampus.
```{r PaS-Scheme, fig.cap = '(ref:PaS-SchemeShortCaption)', fig.scap='(ref:PaS-SchemeCaption)', echo=FALSE, fig.width=8, fig.asp=0.4, fig.show='hold', fig.pos='h'}
horPaS  <- grImport2::pictureGrob(grImport2::readPicture(rawToChar(rsvg::rsvg_svg("Figures/Schemes/PaS_horizontal_clean.svg"))))
sagPaS  <- grImport2::pictureGrob(grImport2::readPicture(rawToChar(rsvg::rsvg_svg("Figures/Schemes/PaS_sagittal_clean.svg"))))
Schematic_PaS_hor_Image_Plot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/Schematic_PaS_hor_Image_Plot.rds")
wrap_elements(gridExtra::arrangeGrob(sagPaS), clip = T) + wrap_elements(gridExtra::arrangeGrob(horPaS), clip=T) + Schematic_PaS_hor_Image_Plot + plot_layout(widths = c(3.2,1.5, 1.4)) & plot_annotation(tag_levels = "A") &
    theme(plot.tag = element_text(size=16))
```


<!-- ## Spatial Navigation -->

<!-- ### Theta -->

