# Materials and Methods {#methods}

## Animals

## Surgical Procedures

### Stereotactic Injections {#injections}
For all surgical procedures mice were anaesthetised with isoflurane (1.5% vol/vol in oxygen, CP-Pharma Handelsgesellschaft mbH, Burgdorf, Germany) and received carprofen (5 $\frac{mg}{kg} \textstyle$) as analgesia additionally to subcutaneous application of lidocaine as local analgesia at the side of incision. Animals were placed into a stereotaxic frame where they were fixed and stabilised with ear bars. After disinfecting the place of incision with iodide (###### find exact one####) an one cm incision was performed on the midline. The skin was pushed and fixed to the side so that the skull was easily accessible.

```{r MS-Injection-Scheme, fig.cap = 'Schematic of medial septum (blue) and the injection canula (black). A 10° angle is used to avoid damage to superior sagittal sinus which is located on the midline on top of the brain surface (not shown).', fig.show='hold', fig.pos='H',echo=FALSE, fig.height=3}
Inj  <- grImport2::pictureGrob(grImport2::readPicture(rawToChar(rsvg::rsvg_svg("Figures/Schemes/MS_schema.svg"))))
wrap_elements(gridExtra::arrangeGrob(Inj), clip=T)
```

The skull was virtually levelled using a robot stereotaxic (Neurostar, Tübingen, Germany) and a craniotomy drilled at the position of entry (in mm from bregma: 0.7 anteriorposterior, 0.7 mediolateral). A NanoFil syringe (World Precision Instruments, LCC., Sarasota County, FL, US) with a 34‐gauge needle with adeno-associated virus 1 (AAV1) containing the plasmid for the yellow fluorescent protein (YFP) and channelrhodopsin (*pAAV-EF1*$\alpha$*-doublefloxed-hChR2(H134R)-EYFP-WPR* at 6.26·10^11^ VG/ml) was placed with a 10° angle through the hole and slowly moved 3.8 mm downwards along the dorsoventral axis to the MS (Figure \@ref(fig:MS-Injection-Scheme)). The position was held for 5 min before 200 nl of the virus was injected at a rate of 40 nl/min. The needle was left for further 5 min until it was slowly withdrawn. During the whole procedure the animal's temperature was maintained at 37°C, eyes were covered with eye-cream to provide moisture, and anaesthesia was checked regularly. The incision was closed and the animal was put back to the home-cage to recover where it woke up after less than 15 min. Animals were checked daily and were allowed to recover for at least 4 weeks.

### Headbar Implant
Previously injected animals as part of *in-vivo* awake head-fixed recordings were given a headbar implant. Surgical procedures were similar to \@ref(injections), however, after the incision a craniotomy was made at a more anterolateral position (in mm from bregma: 1.2 anteriorposterior, 1.2 mediolateral) to place a screw for grounding. The screw was fixed without damaging the surface of the brain. The skull surface was then sealed using OptiBond™ Solo Plus (KerrHawe SA, Bioggio, Switzerland) and the golden pin attached to the screw was stabilised using CHARISMA® ABC A1 (Kulzer GmbH, Hanau, Germany). A headbar was placed over the skull, aligned to midline and fixed with Paladur® dental cement (Kulzer GmbH, Hanau, Germany). The animal was then released back to the home-cage where it could recover for a at least 2 days before habituation began.

## *In-Vitro* Recordings

### Slice Preparation
For slices animals (n = XX PV-Cre mice, n = XX ChAT-Cre mice ) were deeply anaesthetised under isoflurane and then decapitated. The brain was quickly removed and transferred to ice-cold (~4°C) slicing solution (sucrose artificial cerebral spinal fluid -- sACSF). sACSF contained (mM): 87 NaCl, 26 NaHCO~3~, 10 Glucose, 50 Sucrose, 2.5 KCl, 1.25 NaH~2~PO~4~, 0.5 CaCl~2~, 3 MgCl~2~·6H~2~O.
The forebrain was separated from the rest by making a coronal cut. The posterior part of the brain was separated with an sagittal cut in the two hemispheres. The two posterior Hemispheres and the forebrain were placed on a vibrating blade microtome (VT1200S, Leica Biosystems, Wetzlar, Germany) to produce 400 µm horizontal and coronal slices (Figure \@ref(fig:PaS-Scheme) B) which were then transferred to an interface chamber for up to 1-6 h storage. Slices were superfused constantly with oxygenated ACSF (in mM: 119 NaCl, 26 NaHCO~3~, 10 Glucose, 2.5 KCl, 1 NaH~2~PO~4~, 2.5 CaCl~2~, 1.3 MgCl~2~·6H~2~O). All solutions were saturated in oxygen using carbogen (95% O~2~ and 5% CO~2~).
To verify channelrhodopsin expression fluorescence was checked in medial septal coronal sections under a fluorescence microscope (DM3000, Leica Biosystems, Wetzlar, Germany) prior to recordings.

### Slice Recordings
Slices were transferred from the interface storage to the recording chamber where they were constantly superfused (at 3.5 ml/min) with oxygenated ACSF and maintained at 32-34°C. Cells were visually identified using an upright microscope (BX51W1, Olympus, Tokyo,  Japan) and infrared differential interference contrast microscopy through a digital camera (XM10-IR, Olympus, Tokyo, Japan). To record in the the whole-cell patch-clamp configuration a borosilicate glass electrode (Harvard Apparatus, Holliston, Massachusetts, USA) was pulled on a DMZ-Universal-Electrode-Puller (Zeitz-Instruments, Martinsried, Germany) to 2.5-6 M$\Omega$ and filled with intracellular solution (in mM: 120 K-Gluconate, 10 HEPES, 10 KCl, 5 EGTA, 2 MgSO~4~·7H~2~O, 3 MgATP, 1 NaGTP, 5 Phosphocreatine Na, 0.2% Biocytin). Data was recorded with a Multiclamp 700A/B (Molecular Devices, LLC., San Jose, CA, USA) were signals were filtered at 10 kHz, sampled at 20 kHz and then digitised. Data was either digitised by a Digidata 1550 (Molecular Devices, LLC., San Jose, CA, USA) for use in pClamp. In recordings in IGOR Pro 6.12 (WaveMetrics, Inc., Portland , OR, USA) or WinWCP V5.3.7 [@dempster_new_1997] a BNC-2090 interface board (PCI 6035E A/D Board, National Instruments, Austin, Texas, USA) or a USB-6229 BNC (National Instruments, Austin, Texas, USA) was used for digitisation. After opening, cells were held in voltage-clamp at -60 mV to measure series resistance and assess access. Cells were then switched to current-clamp configuration to perform a characterisation of somatic properties (current steps of 40-100 pA). To assess septal connectivity to the parasubiculum channelrhodopsin expressing fibres in the area were activated by 10 ms light pulses from a LED (CoolLED pE-2, Andover, UK) or an halogen lamp (### check for model ####) coupled to the 60x objective. To elicit inhibitory (IPSPs) or excitatory postsynaptic potentials (EPSPs) and inhibitory (IPSCs) or excitatory postsynaptic currents (EPSCs) a range of light pulse frequencies (10, 20, 40 Hz) was used. Additionally, inputs hidden due to low driving force or simultaneously occurring synaptic inputs were unmasked by changing the holding potential of the cell from -60, to -80 and then to -50 mV after at least 10 trials respectively. In some recordings (##### N number #####) monosynaptic inputs were validated by at first blocking with 1 µM tetrodotoxin (TTX) and then reinstating inputs by depolarising the terminals with ##µM## 4-Aminopyridine (4-AP).
Liquid junction potential was not corrected for. Slices were fixed in 4% paraformaldehyde (PFA) overnight for immunohistochemistry stainings. 


Drugs used:
4-Aminopyridine (Tocris Bioscience, Bristol, UK)
NBQX disodium salt (Tocris Bioscience, Bristol, UK)
SR 95531 hydrobromide (Tocris Bioscience, Bristol, UK)

## *In-Vivo* Recordings

### Anaesthesised Recordings
Before recordings the animals was anaesthetised as described under \@ref(injections). A craniotomy was performed at 1 anteroposterior and 1 mediolateral (in mm from bregma) and two additional ones at ±2.5 mediolateral and 0 ####### anteroposterior (in mm from lamda). The animal was injected with ### $\frac{mg}{kg} \textstyle$ urethane and subsequently transferred to the recording setup where two ISO-3x-tet-lin optrodes (NeuroNexus, Ann Arbor, MI, USA) with 32 channels each were used to record in the MS and PaS simultaneously and allowed for light stimulation of both regions. The probe targeting the medial septum was placed with an angle of 10° and lowered 4.3 mm from the surface. The optrode targeting the PaS was lowered perpendicular to the brains surface with 2.5-35 mm distance from the surface. 

### Head-Fixed Awake Immobile Recordings {#immobile}
For head-fixed recordings in immobile mice animals were habituated to the setup and trained to stand still in a paper roll whilst being fixed at the headbar. Animals were trained for 5-7 days with increasing time duration until they were able to remain still for up to an hour. After training sessions mice were rewarded with condense milk. After the training period animals were anaesthetised as described in \@ref(injections). A craniotomy was performed at 1 anteroposterior and 1 mediolateral (in mm from bregma) and another craniotomy at -2.5 mediolateral and 0 ####### anteroposterior (in mm from lamda). The probe targeting the medial septum was placed with an angle of 10° and lowered 4.3 mm from the surface. The optrode targeting the PaS was lowered perpendicular to the brains surface with 2.5-35 mm distance from the surface. After the recording session Kwik-Cast (World Precision Instruments Ltd, Hertfordshire, UK) was applied to the craniotomies to protect the brain from drying out.

### Head-Fixed Awake Recordings on Treadmill
For head-fixed recordings on the treadmill animals were water deprived to encourage them to drink during the training sessions. During the training sessions animals were trained to run on the treadmill whilst being fixed at the headbar. 


```{r LinearTrackScheme, fig.cap = 'Schematic of linear track. In A the wall of the reward zone (120-140 cm) is striped so that it can be distingushed by the mouse from the rest of the environment. The animal had to reach the reward zone, wait for 3 s and then got a water reward. B represents the same task but without the beaconed reward zone.',echo=FALSE, fig.width=6}
#knitr::include_graphics("Figures/Schemes/LinearTrackCombine.pdf")
beaconed<-wrap_elements(cowplot::ggdraw()+cowplot::draw_image("Figures/Schemes/LinearBeaconed.svg"))
nonbeaconed<-wrap_elements(cowplot::ggdraw()+cowplot::draw_image("Figures/Schemes/LinearNonBeaconed.svg"))
beaconed / nonbeaconed & 
  plot_annotation(tag_levels = "A") &
  theme(plot.tag = element_text(size=24))
```

The virtual reality (VR) task entailed a linear track which was 200 cm long and had a reward zone at 120 cm to 140 cm (Figure \@ref(fig:LinearTrackScheme)). The animal firstly, had to reach the reward zone and spent a fixed amount of time (3 s) until a water reward was delivered via a pump. Subsequently, the animal had to reach the end of the track where it was then teleported back to the start. After a waiting period of 5 s a new trial began. Animals were trained for 3 weeks with increasing time duration until they were able to run and perform a distance estimation task. For the recording session animals were surgically prepared as described in \@ref(immobile) and placed on the treadmill to perform the task. Light stimulation was used for a whole trial after a minimum of 5 trials. At the end of the trial light was switched back off and the animal continued with the trials. This alternating procedure was repeated for the whole recording session in order to gather multiple trials during stimulation. During training sessions four beaconed trials (Figure \@ref(fig:LinearTrackScheme)A) were followed by an non-beaconed trial (Figure \@ref(fig:LinearTrackScheme)B) which had no distinguishable reward zone. For Recording sessions the frequency was increased to 2 beaconed trials followed by a non-beaconed trial.

## Histological Processing
Animals from *in-vivo* recordings were perfused after the final recording session. Therefore, they were deeply anaesthetised using urethane (2.5 $\frac{g}{kg} \textstyle$ body weight) and transcardially perfused with phosphate-buffered saline (PBS) followed by 4% PFA. The fixed brain was removed and stored in 4% PFA over night. After post-fixation brains were washed in PBS before being they were dissected into two hemispheres and a forebrain part. The three parts were blocked onto a vibrating blade microtome (Leica VT1000S, Leica Biosystems, Wetzlar, Germany) and sliced into 75 µm thick sagittal (hemispheres containing the PaS) and coronal (forebrain containing the MS) sections.


## Acquisition and Stimulation
Signals were acquired and digitised using a RHD2000 system (Intan Technologies, Los Angeles, CA, USA). The light stimulation was performed using Spike2 coupled to an Power1401-3A data acquisition interface (Cambridge Electronic Design Limited, Cambridge, UK). Custom scripts were used to drive pulse and sine wave stimulations at different frequencies (1 Hz, 2 Hz, 4 Hz, 8 Hz, 16 Hz, 32 Hz). 1 Hz (2 ms pulse length) stimulation was used to identify channelrhodopsin expressing cells in the MS whereas higher frequencies were used to drive network activity in the MS and the PaS. For pulse stimulations between 2 and 8 Hz nested pulse bursts (5 times 5 ms single pulse) where used to imitate theta burst activity. Pulse frequencies at 16 and 32 Hz were achieved through single 10 ms long pulses.

## Immunohistochemistry and Histology
Anti-NeuN purified Antibody (ABN90P, Sigma-Aldrich, St. Louis, MO, United States, Sigma-Aldrich, St. Louis, Missouri, United States)
Anti-Reelin Antibody (MAB5364)
WFS1 Polyclonal antibody - Rabbit/IgG (11558-1-AP, Proteintech Group, Inc, Rosemont, IL, USA)

## Statistical Analysis
