# Results {#results}

```{r CharParFileLoad, echo=F}
CellChar <- readRDS(file = "/alzheimer/Daniel_Data/R/In-Vitro-ConnectivityAnalysis/Output/CellChar.rds")
DSCount <- CellChar[CellTypeLayer=="DS",.N,]
INCount <- CellChar[CellTypeLayer=="IN",.N,]
IN1Count <- CellChar[CellTypeLayer=="IN_1",.N,] 
FSCount <- CellChar[CellTypeLayer=="FS",.N,] 
AHPDiffRange <- as.vector(unlist(ParameterCompTable[Parameter=="AHP (mV) (µ)"&OutOfInterval==1,.(min(`50%`), max(`50%`)),]))
ParameterTable <- readRDS(file = "/alzheimer/Daniel_Data/R/In-Vitro-ConnectivityAnalysis/Output/ParameterTable.rds")
#ParameterCompTable <- readRDS(file = "/alzheimer/Daniel_Data/R/In-Vitro-ConnectivityAnalysis/Output/ParameterComparisons.rds")
```
## *In-Vitro*

### Cell Characterisation

Since there is only limited information about electrophysiological properties of cells in the PaS and none for layer 1 interneurons we thought out to describe the characterisation of a wider range of parameters. To get a good estimate for these parameters it is required to get a large enough sample of cells. We recorded `r CellChar[,.N,]` neurons in the PaS and quantified a total of $19$ cell parameters giving a rich overview of electrophysiological features we hope can be used in the future and help to understand differences in the PaS better. As we described in @sammons_electrophysiological_2019 cells in the PaS can be split into 3 major clusters which correspondent to mainly a split in GABAergic and glutamergic cells. This was determined previously in @sammons_electrophysiological_2019 by recording cells in VGAT^+^ mice. To get similar separation of classes I integrated the measured properties in the same  hierarchical clustering approach we used previously. The cell parameters were firstly scaled and centred by subtracting the mean and dividing by the standard deviation for each parameter respectively. This was then followed by calculating the euclidean distance matrix which was used to cluster the cells based on Ward's method. As demonstrated by @sammons_electrophysiological_2019 using this approach cells in the PaS can be divided into $3$ clusters which represent putative fast spiking interneurons (FS), a heterogeneous cluster with mainly regular interneurons (IN) and pyramidal cell which are often WFS1^+^. For this data set I was additionally interested in differentiating between layer 1 interneurons and other cells. Therefore, I further stratified the cells based on location determined on confocal images or differential interference images at the recording setup. Only $1$ layer 1 cell was originally classified to belong to the pyramidal cell cluster and $34$ were determined to be in the IN cluster. This classification underlines the notion that the IN cluster is dominated by interneurons in the PaS. A substantial amount of the cells were putative pyramidal cells (`r DSCount`) whereas a fraction were classified as interneurons (`r FSCount` Fast-Spiking Basket cell like interneurons (FS); `r INCount` layer 2 interneurons (IN); `r IN1Count` layer 1 interneurons (IN-L1)) making it `r round(CellChar[CellTypeLayer=="DS",.N,]/CellChar[,.N,]*100)`% putative pyramidal cells. We tried to represent other classes equally well which meant a targeted approach was necessary to increase the samples of the IN, FS and IN-L1 clusters. Therefore, we visually identified non-pyramidal cells based on the soma shape and orientation and structure of the dendrites. Cell properties were measured in whole-cell current-clamp configuration using hyper- and depolarisation current steps (step size of $40pA$) (Figure \@ref(fig:CellCharacterisationFigure)A). Here we report for the first time a description of electrophysiological properties of layer 1 cells in the PaS an often overlooked group of cells. Overall cells showed a similar resting potential (`r HDICt(ParameterTable[Parameter=="Vm (mV)",], unit="mV")`) with differences in the input resistance ($R_{in}$) showing lower resistance in FS cells and very high $R_{in}$ measured in IN-L1 cells (`r HDICt(ParameterTable[Parameter=="Rin (MOhm)",], unit="M\\Omega")`) compared to others. The membrane capacitance ($C_{m}$), a proxy for different properties of the cell such as size or conductance, was very low in IN-L1 cells which usually had visually a small soma and shorter dendrites than pyramidal cells or FS basket like cells (`r HDICt(ParameterTable[Parameter=="Cm (pF)",], unit="pF")`). The strongest mean difference between cell types was measured between DS and IN-L1 cells between the different cell types (`r HDICt(ParameterTable[Parameter=="Vm (mV)",], unit="mV")`). This low $C_{m}$ corresponded to the $R_{in}$ except that even compared to cells from the IN cluster IN-L1 cells had a lower $C_{m}$ (`r HDICt(ParameterCompTable[grepl("Cm",Parameter)&Comparison=="IN-L1-IN"], unit = "pF")`) underlining that possibly the compact cell size of IN-L1 cells could be a contributing factor. However, the combination of $C_{m}$  and $R_{in}$ did not result in the highest $firing$ $frequency$ for IN-L1 cells (`r HDICt(ParameterTable[Parameter=="Firing Frequency (Hz)",], unit="Hz")`). In fact, the cell cluster we termed fast spiking interneurons -- FS -- showed the largest mean difference in $firing$ $frequency$ (Figure \@ref(fig:CellCharacterisationFigure)E) compared to all other clusters and could reach frequencies 2-3 times higher than the other groups (`r HDICt(ParameterCompTable[grepl("firing frequency", Parameter, ignore.case = T)&grepl("FS",Comparison),], unit = "Hz")`). In summary, all cells had a higher $firing$ $frequency$ compared to putative glutamergic pyramids (`r HDICt(ParameterCompTable[grepl("firing frequency", Parameter, ignore.case = T)&grepl("DS",Comparison),], unit = "Hz")`). A similar relationship was observed when analysing the $sag$ $ratio$ (Figure \@ref(fig:CellCharacterisationFigure)G). The sag-ratio is a measure of rebound after hyperpolarisation (`r HDICt(ParameterTable[Parameter=="Sag-Ratio",], unit="")`) and was higher in all interneuron classes compared to the DS cluster `r HDICt(ParameterCompTable[grepl("sag", Parameter, ignore.case = T)&grepl(pattern = "DS",Comparison),])`). This finding suggests that rebounds due to hyperpolarisation are more pronounced in glutamergic cells than in interneurons. In contrast, we could not detect reliable differences in the medium after hyperpolarisation (mAHP) -- the corresponding response after a finished depolarisation step (`r HDICt(ParameterTable[Parameter=="mAHP (mV)",], unit="mV")`).

(ref:CellCharacterisationFigureShortCaption) Cell characterisation
(ref:CellCharacterisationFigureCaption) **Cell characterisation.** Characterisation of cells in the PaS after current step injection (A) with the first spike coloured in red and the following in black. Other current injections are in grey. Scale bars reflect 0.1s and 20mV of Amplitude. B-G show the measured cell properties of the 214 cells as point clouds with its median as line and the bar plot represents the population estimates given the cell type.
```{r CellCharacterisationFigure, fig.cap = '(ref:CellCharacterisationFigureCaption)', fig.scap='(ref:CellCharacterisationFigureShortCaption)', fig.height=8, fig.width=6, echo=F, fig.align='center'}
CellCharPropertiesPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/In-Vitro-ConnectivityAnalysis/Output/CellCharPropertiesPlot.rds")
# take mAHP out?
CellCharPropertiesPlot
```

To assess the excitability of cells the rheobase -- the current at which the first spike is trigged --  (Figure \@ref(fig:APPropertyPlot)H) was measured and quantified. DS cells and IN-L1 cells showed the lowest rheobase indicating a high excitability when depolarised `r HDICt(ParameterTable[Parameter=="Rheobase (pA)",], unit="pA")`. This excitability was mainly achieved due to the high $R_{in}$ and not a lower threshold in DS and IN-L1 cells. On the contrary, a lower threshold was measured in FS cells opposed to DS and IN-L1 cells (Figure \@ref(fig:APPropertyPlot)B) underlining the importance of the $R_{in}$ as the possible main mechanism for excitability in the two classes (`r HDICt(ParameterTable[Parameter=="Threshold (mV)",], unit="mV")`). A neuron can have different stages of excitation which can be a result of fast spike initiation, high firing frequency or sustained firing (see adaptation in Table \@ref(tab:CellParTable)). Considering the low rheobase of DS neurons in terms of spike initiation DS cells were overall less excitable than the interneurons when looking at spiking output over different current injections. To quantify this overall excitability I computed the gain for all the neurons (Figure \@ref(fig:APPropertyPlot)I: `r HDICt(ParameterTable[Parameter=="Gain (Spikes/pA)",], unit="\\frac{Spikes}{pA}")`). The gain represents the increase in spiking relative to the input current ($Spikes/pA$) and is substantially higher in interneurons and especially in FS cells (`r HDICt(ParameterCompTable[grepl("gain", Parameter, ignore.case = T),], unit="\\frac{Spikes}{pA}")`). This translates to a more efficient input-output relationship in interneurons compared to DS cells allowing themm to increase spiking activity fast upon depolarisation. Additionally, to the initiation of spiking and the gain, neurons were stimulated as mentioned before with a higher current step to induce the maximum firing frequency. Another different measure of excitability can be the latency of a spike (Figure \@ref(fig:APPropertyPlot)D) measured at the current injection step with the first occurrence of action potentials. This parameter refers to the time it takes for the first spike after the depolarisation onset and reflects how instantaneous or filtered a potential input to a cell is relayed further downstream. Spike latency was on average lower in all interneurons compared to DS cells (`r HDICt(ParameterCompTable[Parameter=="Latency (s) (µ)",], unit="s")`) indicating a fast response upon depolarisation, however, all cell types showed a wide range of possible latencies when spiking. Therefore we termed the putative pyramidal cell cluster "delayed spiking cells" or "DS". This behaviour suggests an interesting property of DS cells where they take time to integrate input current until the threshold is reached. The slower integration is further supported as mentioned before by a higher $C_{m}$ in DS cells (`r HDICt(ParameterCompTable[grepl("Cm",Parameter)&grepl(pattern = "DS", Comparison),], unit = "pF")`).

(ref:APPropertyPlotShortCaption) Action potential properties
(ref:APPropertyPlotCaption) **Action potential properties.** (A) Representative examples of action potentials of the four cell clusters with their properties (B-I). The point clouds in B-I represent measurements of the cells and the corresponding boxplots reflect the estimated cell type population properties.
```{r APPropertyPlot, fig.cap = '(ref:APPropertyPlotCaption)', fig.scap='(ref:APPropertyPlotShortCaption)',echo=F, fig.height=3, fig.asp=1.1, fig.align='center'}
APPropertiesPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/In-Vitro-ConnectivityAnalysis/Output/APPropertyPlot.rds")
APPropertiesPlot

```
To address possible implications for cell outputs which can be due to action potential shape I quantified parameters describing the kinetics of APs (Figure \@ref(fig:APPropertyPlot)A). Several parameters were included in the analysis such as the full width at half maximum (Figure \@ref(fig:APPropertyPlot)F), The rising and falling component of the action potential (positive and negative slope -- see Table \@ref(tab:CellParTable)), and the resulting slope-ratio (Figure \@ref(fig:APPropertyPlot)G). These kinetic measures were often showing a wider and slower action potential in DS cells opposed to faster interneuron kinetics (FWHM -- `r HDICt(ParameterTable[Parameter=="FWHM (ms)",], unit="ms")`; sag-ratio -- `r HDICt(ParameterTable[Parameter=="Sag-Ratio",], unit="")`). Other parameters of the action potential are the action potential height (`r HDICt(ParameterTable[Parameter=="AP-Height (mV)",], unit="mV")`) which showed that on average interneurons had smaller APs than DS cells (`r HDICt(ParameterCompTable[Parameter=="AP Height (mV) (µ)",], unit="mV", sig="all")`) and IN and IN-L1 cells had the smallest APs (Figure \@ref(fig:APPropertyPlot)E). On the other hand, FS and IN cells had compared to DS or IN-L1 cells a larger after hyperpolarisation (AHP) directly followed after the action potential (`r HDICt(ParameterCompTable[Parameter=="AHP (mV) (µ)",], unit="mV")`). FS cells showed an especially prominent AHP after the action potential (Figure \@ref(fig:APPropertyPlot)C and Table \@ref(tab:CellParCompTable))) where in contrast DS cells and IN-L1 showed only little of an hyperpolarising overshoot. A full list of parameter estimates can be found in Table \@ref(tab:CellParTable) of the appendix with all performed comparisons (Table \@ref(tab:CellParCompTable)).


### Cell Connectivity

```{r LoadReversalData, echo=FALSE}
ReversalSlopeValues <- CellTypeReversal  <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ReversalSlopeValues.rds")
#### reversal potential
CellTypeReversal  <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/reversalPotential.rds")
colnames(CellTypeReversal) <- c("FS", "IN")
CellType50CC <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/CellType50CC.rds")
Connectivity_CellType <- readRDS(file = "/alzheimer/Daniel_Data/R/In-Vitro-ConnectivityAnalysis/Output/Connectivity_CellType.rds")
HDILogOddsComparison <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/HDILogOddsComparison.rds")
HDILogOddsComparison[!grepl(pattern = "ChAT", GroupComparison), MouseLine:="PV"][!grepl(pattern = "PV", GroupComparison), MouseLine:="ChAT"]
HDILogOddsComparison[is.na(MouseLine), MouseLine:="Comb"]

```
#### PV^+^ projections to the PaS {#PVConnectivity}
To investigate whether GABAergic PV^+^ MS cells project to the PaS and which targets they have we recorded in whole-cell patch-clamp configuration *in-vitro* in PV-Cre. GABAergic P^+^ cells in the MS were expressing channelrhodopsin when cells were infected using AAV-virus and fibres could be found in the PaS (Figure \@ref(fig:PVConnectivity)A) after at least $4$ weeks of expression. After the cell characterisation we illuminated the septal terminals in the PaS using blue light and recorded inhibitory postsynaptic currents and potentials at different holding potentials ($-50mV$, resting potential ($Vm$), $-80mV$ Figure \@ref(fig:LED10CC)). We stimulated cells over several trials ($>10$) and averaged the signal to increase signal-to-noise for detection. In total we tested $`r CellChar[MouseLine=="PV",.N,]`$ cells and got $`r CellChar[MouseLine=="PV"&Connected==1,.N,]`$ responses which were all hyperpolarising inputs except for one input onto a DS cell which showed an EPSP and could be abolished by NBQX. The $10ms$ stimulation reliably resulted in a fast hyperpolarisation which could be reversed when changing the holding potential (Figure \@ref(fig:InputReversalPlot)A and B).

(ref:LED10CCShortCaption) Light induced synaptic inputs by activation of PV^+^ fibres 
(ref:LED10CCCaption) **Light induced synaptic inputs by activation of PV^+^ fibres .** A shows a FS and an IN cell receiving GABAergic septal inputs. B inhibitory postsynaptic potentials at different holding potentials during 10Hz stimulation elicited with 10ms light pulses. Blue lines indicate the timing and length of stimulation. C shows a FS cell in voltage-clamp configuration held at -45mV during 4Hz, 10Hz, 20Hz, and 40Hz stimulation with 10ms light pulses.
```{r LED10CC, fig.cap='(ref:LED10CCCaption)', fig.scap='(ref:LED10CCShortCaption)',echo=F, message=FALSE, warning=FALSE, fig.width=8, fig.align='center', fig.asp=0.6, fig.pos="H"}

### add resting potential information
LED10CCPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/LightStimulation10Hz.rds")
LED10CCPlot
```

An important feature of synaptic connections is the strength of the connection which can be quantified as amplitude. A large negative deflection will likely be able to inhibit a cell when depolarised. To estimate the reversal potential and the amplitudes of synaptic inputs I used a Bayesian hierarchical linear model. These models have the advantage that accounting for nested structures in the data (cells, animal, experimenter) is simple and will result in more unbiased population parameter estimates. I reparameterised the regular linear regression ($y = m \cdot x + n$) to estimate directly the reversal potential ($y=m \cdot x - slope \cdot V_{rev}$) and added an informative prior (some restricting soft bounds) with mean of $-68mV$ taken from @gonzalez-sulser_gabaergic_2014 and $3$ standard deviations. Since a total of only $`r CellChar[MouseLine=="PV"&Connected==1&CellType=="IN",.N,]`$ interneurons from layer 1 and 2 were connected I pooled them for further analysis. A comparison of the reversal potential showed that there was no detectable difference `r HDICt(HDIcalc(CellTypeReversal), unit="mV")` between FS and IN/IN-L1 clusters and a calculation of the slope difference revealed that even though there might be a trend for IN cells to receive stronger inputs at any given holding potential it did not sufficiently differ (`r HDISummary(HDIcalc(DifferenceMat(ReversalSlopeValues, GroupStrings = c("FS", "IN"))), unit = "mV_{amp}/mV_{hold}")`). This was underlined by just extracting the posterior distribution for $-50mV$ holding potential to investigate differences in strongly depolarised cells (Figure \@ref(fig:InputReversalPlot)C and D). As seen in D the the difference in mean amplitude between cell types was `r HDItxt(CellType50CC)` again showing a trend for IN/IN-L1 cells to have slightly larger IPSPs. Considering that $R_{in}$ and other passive membrane properties differ between the cell types (Table \@ref(tab:CellParCompTable)) it is possible that they could contribute to the variability between FS and IN/IN-L1 clusters. 

(ref:InputReversalPlotShortCaption) Reversal input amplitude
(ref:InputReversalPlotCaption) **Reversal input amplitude.** Amplitudes of GABAergic medial inputs to FS and IN cells of the PaS given a holding potential of a cell (A). Single dots represent a single input and the line is the median estimated population mean conditional on the cell type. The grey shaded area shows the 95\% highest density interval. (B) The estimated reversal potential for FS and IN cells. (C) A subset of inputs where the baseline was between -52mV and -48mV is shown next to the estimated input amplitude given the baseline is 50mV. The line in the point cloud represents the mean of the subset. (D) A comparison of the FS and IN responses reveal a tendency that IN inputs are larger but the zero-line is contained in the 95\% credible interval.
```{r InputReversalPlot, fig.cap='(ref:InputReversalPlotCaption)', fig.scap='(ref:InputReversalPlotShortCaption)',echo=F, message=FALSE, warning=FALSE, fig.height=7, fig.width=6, fig.align='center', fig.pos='H'}
ReversalLinePlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/ReversalLinePlot.rds")
ReversalPotentialPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/ReversalPotentialPlot.rds")
AmpPVAt50mVCC <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/AmpPVAt50mVCC.rds")
CellType50CCDeltaPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/CellType50CCDeltaPlot.rds")

DesignMatrix <- "
AAAAAA
AAAAAA
AAAAAA
BBCCCD
BBCCCD
"
ReversalOverview <- ReversalLinePlot + ReversalPotentialPlot + AmpPVAt50mVCC+ CellType50CCDeltaPlot + plot_annotation(tag_levels = "A") + plot_layout(design = DesignMatrix, nrow = 1) & 
  theme(plot.tag = element_text(size=16))
ReversalOverview
# 
# ReversalLine + ReversalPot + plot_annotation(tag_levels = "A") + plot_layout(widths = c(6,1), ncol = 2) & 
#   theme(plot.tag = element_text(size=16))
```

(ref:GabazineWashinShortCaption) Gabazine washin and input latency
(ref:GabazineWashinCaption) **Gabazine washin and input latency** (A) Traces of light induced IPSPs (top) show that after washin of Gabazine IPSPs are abolished (bottom). (B) Gabazine reduces amplitude of IPSPs to noise level (B). (C) The estimated input latencies (for instantaneous LED light stimulation) with the raw data points for the inputs (FS: $`r paste0(InputOnsetsCells[CellType=="FS",N,],"/",InputOnsets[CellType=="FS",N,])`$; IN: $`r paste0(InputOnsets[CellType=="IN",N,],"/",InputOnsetsCells[CellType=="IN",N,])`$).

```{r GabazinePlot, fig.cap='GabazineWashinCaption', fig.scap='GabazineWashinShortCaption',echo=F, message=FALSE, warning=FALSE, fig.width=7, fig.align='center', fig.height=4, fig.pos='H'}

Onset_CellTypeShutter <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/Onset_CellTypeShutter.rds")*1e3
colnames(Onset_CellTypeShutter) <- c("FS", "IN")
OnsetShutter_HDI <- HDIcalc(Onset_CellTypeShutter)

OnsetLED <- readRDS("Data/Onset_CellTypeLED.rds")*1e3
colnames(OnsetLED) <- c("FS", "IN")
OnsetLED_HDI <- HDIcalc(OnsetLED)

### add GABAzine Model output
### add resting potential information

GabazineTrace <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/FS_GabazinePlot.rds")
GabazineAmpPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/GabazineAmpPlot.rds")
OnsetLEDPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/OnsetLEDPlot.rds")
GabazineMu <- readRDS("Data/GabazineMu.rds")

GabazineTrace + GabazineAmpPlot + OnsetLEDPlot + plot_annotation(tag_levels = "A") + plot_layout(ncol = 3, widths = c(1.5,1, 1)) &
  theme(plot.tag = element_text(size=16))
```


To establish the identity of GABAergic inputs we washed in $`r InputParameters[Mouseline=="PV"&Drug==T&!grepl(pattern = "NBQX", Comment),.(unique(Cell)),][,.N,]`$ recordings of FS cells $1\mu M$ Gabazine, a GABA~A~ blocker. Indeed we could show that the application of the drug abolished inputs (Figure \@ref(fig:GabazinePlot)A) and reduced the mean amplitude from `r round(GabazineMu[GroupComparison=="Control",3,],digits=2)`mV `r HDIEx(GabazineMu[GroupComparison=="Control",])` in the control to noise level (`r HDISummary(GabazineMu[GroupComparison=="Gabazine"], unit = "mV")`) as shown in Figure \@ref(fig:GabazinePlot)B. This indicates that septal PV^+^ projections which elicit IPSPs are mediated by GABA~A~-receptor activation. To check whether responses were directly associated with the stimulation I analysed the input latency. In fact, we did not observe slow responses indicating GABA~b~ mediated hyperpolarisations. However, cells in the PaS as in many other areas have GABA~b~ receptors and will hyperpolarise when washing in the GABA~b~ antagonist baclofen (data not shown). For analysis of the latency (Figure \@ref(fig:GabazinePlot)C) I performed a linear hierarchical model with log-normal error distribution which is often observed for timing events. The hierarchical structure allowed me to account for different aspects of the experimental design such as cell types and LED or shutter stimulation. The LED, based on a pure electric activation, will give an instantaneous maximum light power while the shutter will take longer to open completely. This introduces some artefacts which can be accounted for. I chose inputs measured in current-clamp with an amplitude of more than $0.2mV$ and $3$ times the baseline standard deviation to estimate the cell type specific onsets. As prior for the mean I relied on data from @gonzalez-sulser_gabaergic_2014 where in the MEC a mean latency of $1.54ms$ was observed. On top of prior latency I added a larger error ($\sigma$) to account for a large range of values (roughly $0-30ms$). The result from this analysis show that FS cells and IN/IN-L1 cells show very similar latencies (`r HDISummary(OnsetLED_HDI)`) but also revealed that indeed a physical shutter can almost double the times showing the importance of a thorough analysis (`r HDISummary(OnsetShutter_HDI)`). The analysis encouragingly captured the variance and helped to determine that inputs are likely direct.

(ref:PVConnectivityShortCaption) Connectivity of PV^+^ MS cells to PaS 
(ref:PVConnectivityCaption) **Connectivity of PV^+^ MS cells to PaS .** (A) A horizontal section stained for WFS1 (top) and against YFP (bottom) which is expressed by PV^+^ fibres coming from the MS. (B) The estimated connectivity as posterior distribution is shown in grey for different cell types (from top to bottom: DS, FS, IN IN-L1). The dark grey area represents the 95\% credible interval of the posterior distribution and the light grey show the remaining tails. The vertical lines show the median of the posterior and the black dots at the bottom of each distribution the raw frequentist point estimates. (C) Probability distribution can be compared to each other by calculating the log odds ratio for any given comparison. The red line indicates the region of practical equivalence. 

```{r PVConnectivity, fig.cap='(ref:PVConnectivityCaption)', fig.scap='(ref:PVConnectivityShortCaption)',echo=F, message=FALSE, warning=FALSE, fig.width=8, fig.align='center', fig.asp=0.55, fig.pos='H'}
PV_connectivity <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PV_connectivity.rds")
PV_comp <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PV_comp.rds")
PV_PaS_hor_Image_Plot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PV_PaS_hor_Image_Plot.rds")

PV_PaS_hor_Image_Plot + PV_connectivity + PV_comp + plot_annotation(tag_levels = "A") + plot_layout(ncol = 3, width = c(1.7, 1.5,3)) &
  theme(plot.tag = element_text(size=16))
```

To understand the implication of the inhibition for the PaS as a network I quantified the proportions of cells which receive input and calculated the connectivity as a fraction given the cell type. For the quantification of connectivity I used a logistic regression with hierarchical structure which allowed to estimate an overall septal connectivity to then further stratify and differentiate between PV^+^ and ChAT^+^ projections. Additionally, possible errors due to the nested structure of the experiments (cells, animal, and experimenter). A Bayesian approach is a great way to quantify binary data outcomes and was to some degree used in @desikan_target_2018 where a $beta$ distribution was used to estimate credible intervals for connectivity. In our case we will use the continuous $beta$ distribution as well, however, I was also able perform more calculations on the posterior connectivity distribution since I modelled the experiment with its experimental structure. As weakly informative priors for the mean connectivity I used for PV^+^ connections data from @fuchs_local_2016 and @gonzalez-sulser_gabaergic_2014 and added added wide enough errors to allow the data to inform the analysis. For DS cells only $1$ of $`r CellChar[MouseLine=="PV"&CellType=="DS",.N,]`$ cells responded to stimulation which however could be blocked by NBQX. This translates to a credible interval of `r HDItxt(x = Connectivity_CellType[GroupComparison=="PV-DS"])` (Figure \@ref(fig:PVConnectivity)B) indicating that there is little evidence in the data to assume a probability higher than 8% as an upper bound. In contrast a high proportion of FS cells ($`r CellChar[MouseLine=="PV"&CellType=="FS"&Connected==1,.N,]`$ of $`r CellChar[MouseLine=="PV"&CellType=="FS",.N,]`$) received GABAergic inputs. They represent by far the most likely target of PV^+^ MS projections to the PaS (`r HDItxt(x = Connectivity_CellType[GroupComparison=="PV-FS"])`). This high connectivity is in line with previous studies and other areas resulting in disinhibition of the network. Due to a smaller sample size ($`r CellChar[MouseLine=="PV"&CellTypeLayer=="IN"&Connected==1,.N,]`$ of $`r CellChar[MouseLine=="PV"&CellTypeLayer=="IN",.N,]`$) estimates for IN cells resulted in a wider credible interval with `r HDItxt(x = Connectivity_CellType[GroupComparison=="PV-IN"])`. Surprisingly IN-L1 cells, which were previously not reported to receive PV^+^ inputs from the septum, were inhibited through channelrhodopsin stimulation ($`r CellChar[MouseLine=="PV"&CellTypeLayer=="IN_1"&Connected==1,.N,]`$ of $`r CellChar[MouseLine=="PV"&CellTypeLayer=="IN_1",.N,]`$). This means that IN-L1 are receiving GABAergic inputs in `r HDItxt(x = Connectivity_CellType[GroupComparison=="PV-IN-L1"])` of the cases. To compare connectivities of each cell type to each other posterior distributions were used to calculate the log odds ratio (see equation \@ref(eq:logOddsRatio) for LOR). 

\begin{align}
Odds_{p} &= \frac{p}{1-p} \\
LOR_{p_{1},p_{2}} &= \log \frac{Odds_{p_{1}}}{Odds_{p_{1}}}
(\#eq:logOddsRatio)
\end{align}

The log odds ratio has the advantage of being firstly symmetric and secondly a nice interpretation when comparing probabilities since it can be transformed to the odds ratio indicating how much more something is likely to happen over another event. Therefore, a LOR of 0 (OR of 1) would mean that both cell types have an equal chance of receiving PV^+^ inputs. In Figure \@ref(fig:PVConnectivity)B we can see that the strongest difference can be found between FS and DS cells with a LOR of `r HDItxt(HDILogOddsComparison[MouseLine=="PV"&GroupComparison=="PVFS-PVDS",])` meaning the odds of an FS being connected is almost $200$ times higher than for DS cells. The $95\%$ credible interval does not cross $0$ nor does it enter the region of practical equivalence (for equation of ROPE see \@ref(eq:ROPE)) [@kruschke_doing_2015] as indicated by the red line. The ROPE indicates the which LOR values would likely represent equivalence between the compared groups. Similarly strong was the difference between IN-L1 cells and FS cells (`r HDItxt(HDILogOddsComparison[MouseLine=="PV"&GroupComparison=="PVIN_1-PVFS",])`) and moderately strong for IN and DS cells (`r HDItxt(HDILogOddsComparison[MouseLine=="PV"&GroupComparison=="PVIN-PVDS",])`) and IN and FS cells (`r HDItxt(HDILogOddsComparison[MouseLine=="PV"&GroupComparison=="PVIN-PVFS",])`). The LOR for DS to IN-L1 cells was comparable as was the LOR for IN to IN-L1 indicated by the credible interval crossing the ROPE (Figure \@ref(fig:PVConnectivity)B). However, in total this indicates that interneurons are more likely to be inhibited than putative glutamergic DS cells.

\begin{equation}
ROPE = \pi/\sqrt{3} \cdot 0.1
(\#eq:ROPE)
\end{equation}


Given that fast GABA~A~ responses would allow a high time accuracy due to their fast kinetics, PV^+^ cells of the septum therefore could be crucial for timed inhibition of PaS cells. To test for this possibility I depolarised a small subset of cells which received GABAergic inputs so that they were slightly over the spiking threshold resulting in continuous spiking of the cell (Figure \@ref(fig:PhaseLockInVitroPlot)A). The cell was inhibited over several trials (Figure \@ref(fig:PhaseLockInVitroPlot)B) and showed that spikes started reliably at similar times over different trials and fire periodically afterwards until inhibited again (Figure \@ref(fig:PhaseLockInVitroPlot)B and C). Considering this timing of spikes we can represent every spike train for each single trial as a wave where a spike represents the peak of the wave and the middle of the pause the trough. This reparameterisation of a spike train to a sinusoidal wave allows us to calculate the phase at any given time point ($t$) for every single trial. Since a phase can be represented on a unit circle with $x=\cos \theta$ and $y=\sin \theta$ we can for every time point calculate the average vector length $\rho$ over all the trials for any given point in time (see equation \@ref(eq:rho)).

\begin{equation}
\rho_{t}=\sqrt{(\frac{1}{trial_{N}} \sum_{n=1}^{trial_{N}} sin \theta_{t})^2 + (\frac{1}{trial_{N}} \sum_{n=1}^{trial_{N}} cos \theta_{t})^2 }
(\#eq:rho)
\end{equation}

A $\rho$ value of $1$ indicates a perfect alignment whereas a value of $0$ indicates none. In our case a high value translates to a high alignment in spike timing (Figure \@ref(fig:PhaseLockInVitroPlot)D) over all the trials. An increase can be observed close to the stimulation onset and is carried on as also seen in Figure \@ref(fig:PhaseLockInVitroPlot)C for some time before it is lost and $\rho$ returns to a lower value close to the expectation of a random distribution ($\rho_{random}=\frac{1}{\sqrt{trial_{N}}}$). Over all this illustrates that even when a cell is spiking GABAergic inputs from the MS can be sufficient to inhibit the cell enough to lock firing for a brief time. This observation is crucial for efficient phase locking in an active network.

(ref:PhaseLockInVitroPlotShortCaption) Spike phase reset *in-vitro*
(ref:PhaseLockInVitroPlotCaption) **Spike phase rest ** *in-vitro***.** A depolarised spiking putative interneuron (A) is inhibited using 10ms long light pulses at 10Hz. The raster plot (B) shows spike times throughout 20 trials over a time of 0.5s. C ,a closeup of (A), shows complete inhibition of spikes during stimulation times. D shows $\rho$ a value for phase locking of spikes. The maximum value of $\rho$ is 1 assuming perfect phase lock and the chance level 0.22. Blue lines indicate the timing and length of stimulation in A-D.

```{r PhaseLockInVitroPlot, fig.cap='(ref:PhaseLockInVitroPlotCaption)', fig.scap='(ref:PhaseLockInVitroPlotShortCaption)', echo=F, message=FALSE, warning=FALSE, fig.width=8, fig.align='center', fig.asp=0.6, fig.pos='H'}
PhaseLockInVitro <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PhaseLockPanel.rds")
ChatInputCheck <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ChatInputCheck.rds")
ChATDetails <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ChATDetails.rds")
DeflectionTypeCount <- ChATDetails[Type!="indirect",.N,by=.(Type, CellTypeLayer)]

PhaseLockInVitro
```

#### ChAT^+^ projections to the PaS

The second major group of projections from the MS to other regions comes from ChAT^+^ MS cells. Since there is no data on cholinergic septal projections to the PaS we investigated functional connectivity. Similarly, to the test of connectivity in PV-Cre animals we investigated projections from cholinergic MS cells to the PaS. Cholinergic projections are much more challenging to detect compared to glutamergic or GABAergic inputs. The reason for this is firstly the often found smaller amplitude and secondly the heterogeneity of possible membrane responses (\@ref(fig:ChATLEDStimulation)D). 

(ref:ChATLEDStimulationShortCaption) Light induced synaptic inputs by activation of ChAT^+^ fibres 
(ref:ChATLEDStimulationCaption) **Light induced synaptic inputs by activation of ChAT^+^ fibres .** (A) A horizontal section stained for WFS1 (top) and against YFP (bottom) which is expressed by ChAT^+^ fibres coming from the MS. (B) The estimated connectivity as posterior distribution is shown in grey for different cell types (from top to bottom: DS, FS, IN IN-L1). The dark grey area represents the 95\% credible interval of the posterior distribution and the light grey show the remaining tails. The vertical lines show the median of the posterior and the black dots at the bottom of each distribution the raw frequentist point estimates. (C) Probability distribution can be compared to each other by calculating the log odds ratio for any given comparison. The red line indicates the region of practical equivalence.

```{r ChATLEDStimulation, fig.cap='(ref:ChATLEDStimulationCaption)', fig.scap='(ref:ChATLEDStimulationShortCaption)',echo=F, message=FALSE, warning=FALSE, fig.width=8, fig.align='center', fig.asp=0.5}
ChAT_IN1_Char <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_IN1_Char.rds")
ChAT_IN1_Input10CC <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_IN1_Input10CC.rds")
ChAT_IN1_CC_Frequency_Plot <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_IN1_CC_Frequency_Plot.rds")
ChAT_HyperpolarisationType_Plot <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_HyperpolarisationType_Plot.rds")
ChAT_IN1_Overview <- ChAT_IN1_Char + ChAT_IN1_Input10CC + ChAT_IN1_CC_Frequency_Plot +  ChAT_HyperpolarisationType_Plot + plot_annotation(tag_levels = "A") + plot_layout(ncol = 4, widths = c(1.3,1,1.1, 1.3)) & 
  theme(plot.tag = element_text(size=16))
ChAT_IN1_Overview
```
GABAergic responses could also be observed in PaS cells ($1$ in $`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="FS",.N,]`$ FS, $1$ in $`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN",.N,]`$ IN, and $1$ in $`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="DS",.N,]`$ DS) by stimulating ChAT^+^ fibres supporting previous evidence by @desikan_target_2018 that the medial septal ChAT^+^ fibres can co-release GABA. However, compared to PV^+^ projections GABAergic inputs account only for a small proportion of the projection effect ($3$ in `rCellChar[MouseLine=="ChAT",.N,]` cells). We found a total of $`r CellChar[MouseLine=="ChAT"&Connected==1,.N,]-1`$ in $`r CellChar[MouseLine=="ChAT",.N,]`$ cells responding to stimulation of septal fibres. However, the responses were heterogenous with $DeflectionTypeCount[Type=="depol",sum(N),]$/$`r DeflectionTypeCount[,sum(N),]`$ depolarising (FS: $`r DeflectionTypeCount[Type=="depol"&CellTypeLayer=="FS",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="FS",.N,]`$; IN-L1: $`r DeflectionTypeCount[Type=="depol"&CellTypeLayer=="IN-L1",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN_1",.N,]`$; IN: $`r DeflectionTypeCount[Type=="depol"&CellTypeLayer=="IN",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN",.N,]`$; DS: $`r DeflectionTypeCount[Type=="depol"&CellTypeLayer=="DS",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="DS",.N,]`$), $12/24$ connections were hyperpolarising (FS: $`r DeflectionTypeCount[Type=="hyperpol"&CellTypeLayer=="FS",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="FS",.N,]`$; IN-L1: $`r DeflectionTypeCount[Type=="hyperpol"&CellTypeLayer=="IN-L1",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN_1",.N,]`$; IN: $`r DeflectionTypeCount[Type=="hyperpol"&CellTypeLayer=="IN",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN",.N,]`$; DS: $`r DeflectionTypeCount[Type=="hyperpol"&CellTypeLayer=="DS",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="DS",.N,]`$) showing a mixture of depolarisation followed by a hyperpolarisation $2/24$ (FS: $`r DeflectionTypeCount[Type=="comb"&CellTypeLayer=="FS",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="FS",.N,]`$; IN-L1: $`r DeflectionTypeCount[Type=="comb"&CellTypeLayer=="IN_1",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN_1",.N,]`$; IN: $`r DeflectionTypeCount[Type=="comb"&CellTypeLayer=="IN",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN",.N,]`$; DS: $`r DeflectionTypeCount[Type=="comb"&CellTypeLayer=="DS",sum(N),]`$/$`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="DS",.N,]`$). To quantify the overall connectivity of ChAT^+^ fibres we estimated the connectivity with the same model described in the previous section \@ref(PVConnectivity). 

(ref:ChATConnectivityShortCaption) Connectivity of ChAT^+^ MS cells to PaS 
(ref:ChATConnectivityCaption) **Connectivity of ChAT^+^ MS cells to PaS .** (A) A horizontal section stained for WFS1 (top) and against YFP (bottom) which is expressed by ChAT^+^ fibres coming from the MS. (B) The estimated connectivity as posterior distribution is shown in grey for different cell types (from top to bottom: DS, FS, IN IN-L1). The dark grey area represents the 95\% credible interval of the posterior distribution and the light grey show the remaining tails. The vertical lines show the median of the posterior and the black dots at the bottom of each distribution the raw frequentist point estimates. (C) Probability distribution can be compared to each other by calculating the log odds ratio for any given comparison. The red line indicates the region of practical equivalence  


```{r ChATConnectivity, fig.cap='(ref:ChATConnectivityCaption)', fig.scap='(ref:ChATConnectivityShortCaption)',echo=F, message=FALSE, warning=FALSE, fig.width=8, fig.align='center', fig.asp=0.55}

### add resting potential information
ChAT_connectivity <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_connectivity.rds")
ChAT_comp <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_comp.rds")
ChAT_PaS_hor_Image_Plot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/ChAT_PaS_hor_Image_Plot.rds")

ChAT_PaS_hor_Image_Plot + ChAT_connectivity + ChAT_comp + plot_annotation(tag_levels = "A") + plot_layout(ncol = 3, width = c(1.7, 1.5,3)) &
  theme(plot.tag = element_text(size=16))
```
The projections to the PaS were less heterogeneously distributed compared to the fibres in PV-mice (\@ref(fig:PVConnectivity)B) and show dominant projections to the superficial layers (\@ref(fig:PVConnectivity)A) with substantial connections to IN-L1 cells ($`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN_1"&Connected==1,.N,]`$ of $`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN_1",.N,]`$) resulting in an estimated connectivity of `r HDItxt(x = Connectivity_CellType[GroupComparison=="ChAT-IN-L1"])`. Similarly IN were innervated by ChAT^+^ projections ($`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN"&Connected==1,.N,]`$ of $`r CellChar[MouseLine=="ChAT"&CellTypeLayer=="IN",.N,]`$) and showed a similar proportion of cells responding to light stimulation (`r HDItxt(x = Connectivity_CellType[GroupComparison=="ChAT-IN"])`). In contrast FS cells were less likely to receive inputs from MS cholinergic cells compared to MS PV^+^ projections (`r HDItxt(x = HDILogOddsComparison[GroupComparison=="PVFS-ChATFS"])`) but still had a tendency with `r HDItxt(x = Connectivity_CellType[GroupComparison=="ChAT-FS"])`to be stronger connected than DS cells (Connectivity of: `r HDItxt(x = Connectivity_CellType[GroupComparison=="ChAT-DS"])` and LOR of FS to DS: `r HDItxt(x = HDILogOddsComparison[GroupComparison=="PVFS-ChATFS"])`). Even though IN-L1 cells only represent a very small proportion of the cells in the PaS they were overrepresented in terms of cholinergic inputs depolarising inputs. A drive of cholinergic cells in the MS therefore will lead to depolarised IN-L1 cells in the PaS. In contrast projections to FS while being comparatively as probable will likely hyperpolarise FS cells. DS cells are one of groups with the most mixed input types and the lowest connectivity. 


## In-Vivo
Because of the strong GABAergic projections targeting a large proportion of interneurons in the PaS which in return inhibit other local cells in a timely manner I thought out to investigate the consequences on the local network *in-vivo*. Compared to slices the whole network is intact and can still interact on the dorsal-ventral axis. Therefore I thought it was crucial to measure whether this accurate and targeted inhibition we observed *in-vitro* will effectively entrain $\theta$ in the PaS. To test his hypothesis I designed an experiment in head-fixed awake mice and recorded simultaneously in the MS and the PaS with two separate silicon probes which were both fitted with a glass fibre attached to an LED to illuminate both regions respectively. The challenging part of these recordings is the location of either regions. Both are fairly deep and can not be directly accessed from the top without implementing an angle. Once the target is reached I waited for 15 minutes to allow the tissue to readjust and then started recording. After establishing a baseline I started to stimulate the MS and PaS with the LED respectively using a pulse train stimulation at $2Hz$, $4Hz$, $8Hz$, $16Hz$, and $32Hz$ to entrain the network. To identify ChR2-positive neurons in the MS I applied $200$-$300$ $1Hz$ $2ms$ pulses. Since background activity during the recordings is dominated by lower frequencies and frequencies in the $\theta$-band (Figure \@ref(fig:OverViewInVivo)C and D) I repeated stimulation protocols in a pseudo random manner up to $20$ times and applied a frequency normalisation ($1\cdot f$) which removes dominating low frequency activity (Figure \@ref(fig:OverViewInVivo)E). To reduce a potential deactivation of the ChR2 due to over stimulation I switched between MS stimulation and fibre stimulation in the PaS. Since the MEC and the PaS are in close vicinity recordings from both areas could be achieved in some sessions ($4/16$) which offered a chance to directly compare the power and phase locking behaviour in the two structures. This is especially interesting due to the fact that they both get innervated by the MS, receive inputs via the hippocampal loop and project to each other.


(ref:OverViewInVivoShortCaption) Overview of *in-vivo* recordings
(ref:OverViewInVivoCaption) **Overview of** *in-vivo* **recordings.** (A) shows a coronal section of the MS with the tracks of the silicon-probe (top) and YFP expression in the injection site of an PV-Cre animal. (B) A WFS1 labelled sagittal section of the PaS (top) with the visible track of the second silicon-probe (middle) ends in the PaS where the PV^+^ fibres terminate (bottom). (C) A 15s long single trial with a 5s long 8Hz pulse stimulation indicated with a blue line starting after 5s. (D) The resulting power calculated from the wavelet transform of one trial. (E) The power density is derived from the same wavelet transform but uses a frequency normalisation ($1\cdot f$). 
```{r OverViewInVivo, fig.cap='(ref:OverViewInVivoCaption)', fig.scap='(ref:OverViewInVivoShortCaption)', echo=F, fig.height=7, fig.width=6, warning=FALSE}
WTSingleExampleOverviewPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/WTSingleExampleOverviewPlot.rds")
WTSingleExampleOverviewPlot
```

The MS stimulation did drastically increase the power density in the PaS but also in the MEC (Figure \@ref(fig:WaveletPowerNested)). The increase was especially strong in the frequencies $8Hz$, $16Hz$, and $32Hz$ and less pronounced during the $2Hz$ and $4Hz$ stimulation (Figure \@ref(fig:WaveletPowerNested) and \@ref(fig:PSDSingleStimLineFacetPlot)). In contrast to MS stimulation a stimulation of the fibres in the PaS and MEC did not change the power in any of the frequencies. This observation has crucial implications given that a failure to entrain the PaS or MEC using fibre stimulation could hint to higher importance of global MS interactions. To quantify the dependency on the frequency and location of the stimulation I estimated the power using a hierarchical model of the $Gamma$ family.


\begin{align}
Gamma&(\alpha,\beta)\\
\mu &= \frac{\alpha}{\beta}\\
\sigma^{2} &= \frac {\alpha }{\beta ^{2}}
(\#eq:GammaDistribution)
\end{align}

Since power is limited to only positive values and has a longer tail a $Gamma$ distribution \@ref(eq:GammaDistribution) can capture the properties well. Further the model was set up to estimate correlations in the parameters and account for nested structures. Considering the non-linearity which can be observed in Figure \@ref(fig:PSDSingleStimLineFacetPlot) in all conditions I had to consider to include this behaviour in the estimation. Power independently of a ChR2 stimulation or not does change with frequency. Therefore I decided to address this trend by using an asymptotic regression which defines power at $0Hz$ to be $0$ (see equation \@ref(eq:AsymPower)). At the same time it assumes that there is a ceiling and power is not infinite. By using the formula to estimate $\mu_{Power Density}$ with additional predictors such as the stimulation we estimate the effect on $beta_{slope}$ or $beta_{max}$ indicating the ceiling for the power density.

\begin{equation}
\mu_{Power Density} = \beta_{max}-({\beta_{max}-0})\cdot e^{(-\beta_{Slope}\cdot Frequency)}
(\#eq:AsymPower)
\end{equation}

The asymptotic regressing will take care of the limits which are defined by the network and how much power can be generated and the $Gamma$ error distribution with a log-link will assure positive values close to $0$.


(ref:WaveletPowerNestedShortCaption) Power density in the PaS and MEC
(ref:WaveletPowerNestedCaption) **Power density in the PaS and MEC.** Power density of 15s averaged over multiple trials. The y-axis shows the frequency (Hz) and the x-axis the time. Stimulation occurs from 5s to 10s (blue line). Major columns refer to the recording location (PaS or MEC) and minor columns to the stimulation location (MS or fibre stimulation). Each row shows a different stimulation frequency (2Hz, 4Hz, 8Hz, 16Hz, 32Hz). Bright yellow indicates high power density and black low power density. An increase in power can be observed in stimulation blocks at the stimulation frequency when stimulating in the MS and is absent during fibre stimulation.

```{r WaveletPowerNested, fig.cap='(ref:WaveletPowerNestedCaption)', fig.scap='(ref:WaveletPowerNestedShortCaption)', echo=F, fig.align='center', fig.height=5, fig.width=7, warning=FALSE}
PSDWaveletOverview <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PSDWaveletOverview.rds")
PSDWaveletOverview
```

(ref:PSDnestedOverviewShortCaption) Power density spectrum in the PaS and MEC
(ref:PSDnestedOverviewCaption) **Power density spectrum in the PaS and MEC.** Power density spectrum during different stimulation frequencies (rows) extracted from the wavelet transform for time blocks before the stimulation (Pre - black line), during the stimulation (Stim - blue line) and directly after (Post - grey line). The dashed lines indicate the stimulation frequency whereas continuous lines show the mean power density and the SEM. The two left columns show responses in the PaS after septal stimulation (column 1) and stimulation of the fibres in the PaS (column 2). Analogous the two right columns show the response in the MEC. The frequency on the x-axis is log~2~ transformed. PaS with MS stimulation: 9 session/4 animals; PaS with fibre stimulation: 9 session/4 animals; MEC with MS stimulation: 7 session/4 animals; MEC with fibre stimulation: 11 session/6 animals.

```{r PowerDensityNested, fig.cap='(ref:PSDnestedOverviewCaption)', fig.scap='(ref:PSDnestedOverviewShortCaption)', echo=F, fig.align='center', fig.height=6, fig.width=7, warning=FALSE}
PowerDensitySpectralPlotNestedOverview <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PowerDensitySpectralPlotNestedOverview.rds")
PSDN <- readRDS("Data/PSDN.rds")
PowerDensitySpectralPlotNestedOverview
```

(ref:PSDSingleStimLineFacetPlotShortCaption) Power density is increased during MS stimulation
(ref:PSDSingleStimLineFacetPlotCaption) **Power density is increased during MS stimulation.** Power density before (Pre), during (Stim), and after stimulation (Post) of the MS cells somatically (left) and fibre stimulation at the recording site (right). Increased power density during stimulation in the PaS (top) and the MEC (bottom) can be observed with MS stimulation but not using fibre stimulation. PaS with MS stimulation: 9 session/4 animals; PaS with fibre stimulation: 9 session/4 animals; MEC with MS stimulation: 7 session/4 animals; MEC with fibre stimulation: 11 session/6 animals.

```{r PSDSingleStimLineFacetPlot, fig.cap='(ref:PSDSingleStimLineFacetPlotCaption)', fig.scap='(ref:PSDSingleStimLineFacetPlotShortCaption)', echo=F, fig.align='center', fig.height=5, fig.width=7, warning=FALSE}
PSDStimLinelog10PlotFacet <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PSDStimLinelog10PlotFacet.rds")
HDIBetaMaxSummary <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/HDIBetaMaxSummary.rds")
HDIBetaMaxSummaryStimMS <- HDIBetaMaxSummary[grepl(pattern = "MS",GroupComparison)&grepl(pattern = "Stim|Pre",GroupComparison),][,GroupComparison:=gsub(pattern = "-MS", replacement = "", x = GroupComparison),]
HDIBetaMaxSummaryStimFibre <- HDIBetaMaxSummary[grepl(pattern = "fibre",GroupComparison)&grepl(pattern = "Stim|Pre",GroupComparison),][,GroupComparison:=gsub(pattern = "-fibre", replacement = "", x = GroupComparison),]
HDIBetaMaxDiff <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/HDIBetaMaxDiff.rds")
HDIBetaMaxDiffSummary <- HDIBetaMaxDiff[c(53,58),][,GroupComparison:=ifelse(grepl("PaS",GroupComparison), "PaS", "MEC"),]
HDIBetaSlopeSummary <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/HDIBetaSlopeSummary.rds")
HDIBetaSlopeDiff <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/HDIBetaSlopeDiff.rds")
PSDRatio <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PSDRatio.rds")
PSDRatio[,GroupComparison:=RecLoc,]

PSDStimLinelog10PlotFacet

```


As expected the analysis confirmed that the stimulation had an effect on the maximum power in trials where the MS was stimulated directly (`r HDISummary(HDIBetaMaxSummaryStimMS)`). However, a stimulation of only the fibres did not show such a change (`r HDISummary(HDIBetaMaxSummaryStimMS)`). In fact, when calculating the absolute difference between pre and stimulation phases the difference was centred around $0$ (`r HDISummary(HDIBetaMaxDiffSummary)`). This indicates a complete failure of entrainment when only PV^+^ fibres try to drive the LFP in the PaS or MEC. I then calculated the estimated fold change of power density given a frequency (Figure \@ref(fig:PSDModelFrequency)B). Since we estimated the complete posterior distribution of the power density for the pre-stimulation block and stimulation block I could calculate the ratio ($Power_{Stim}/Power_{Pre}$) and compare the effect of the stimulation area (MS Stimulation and fibre stimulation) and frequency. A clear fold change could be observed in the MS stimulation paradigm starting early on (MEC at $3Hz$: `r HDItxt(PSDRatio[OutOfInterval==1][1,])` and PaS at $4Hz:$ `r HDItxt(PSDRatio[OutOfInterval==1][4,])`). This ratio will change with higher frequencies and can reach more than two times the power density of the pre-stimulation block with $32Hz$ stimulation (`r HDISummary(PSDRatio[OutOfInterval==1][(.N):(.N-1),])`). A higher change with higher frequencies could possibly be explained by the generally lower content of absolute power in higher frequencies (Figure \@ref(figOverViewInVivo)D) and therefore can result in larger ratios. On the other hand looking at the power density ratio between pre- and stimulation blocks during fibre stimulation it is apparent that a stimulation effect on the power was absent (bottom Figure \@ref(fig:PSDModelFrequency)B). The credible intervals over all frequencies included the ratio $1$ which means that the two blocks are equal. This can be observed in the PaS and the MEC respectively suggesting similar behaviour of the LFP and possibly similar mechanisms.


(ref:PSDModelFrequencyShortCaption) Power density dependent on stimulation frequency
(ref:PSDModelFrequencyCaption) $\rho$ **Power density on stimulation frequency.** Log Odds Ratio of $\rho$ indicating phase lock during different stimulation frequencies. Black represents measurements before stimulation (Pre), blue during stimulation (Stim), and grey after stimulation (Post). The left half shows responses in the PaS and the right half in the MEC. The top row were measurements from MS stimulation and the bottom row from fibre stimulation in the PaS or MEC. Single dots are values from sessions. The line is the estimated mean population response with the 95\% credible interval indicated as shaded area. The frequency on the x-axis is log~2~ transformed. PaS with MS stimulation: 9 session/4 animals; PaS with fibre stimulation: 9 session/4 animals; MEC with MS stimulation: 7 session/4 animals; MEC with fibre stimulation: 11 session/6 animals.

```{r PSDModelFrequency, fig.cap='(ref:RhoLogOddsModelFrequencyCaption)', fig.scap='(ref:RhoLogOddsModelFrequencyShortCaption)', echo=F, fig.align='center', fig.height=5, fig.width=10, warning=FALSE}
PSDModelFrequencyPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PSDModelFrequencyPlot.rds")
PSDRatioModelFrequencyPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/PSDRatioModelFrequencyPlot.rds")
PSDModelFrequencyPlot + PSDRatioModelFrequencyPlot + plot_layout(widths = c(3,1)) + plot_annotation(tag_levels = "A")  &
  theme(plot.tag = element_text(size=16))

```

Given that I observed an increase in power in the higher frequencies during MS stimulation but no differences when stimulating the fibres locally the question arises whether fibres are more important to set the phase of the LFP than modulating the power The phase represents the time component of periodic signals whereas the power represents the strength and amplitude. I calculated for every single block the wavelet transform $C\psi$ as done before for power calculations but compute the phase using $atan2(\Im(C\psi), \Re(C\psi))$. Instead of averaging the power over trials I calculated the $\rho$ value for any given point in time and any frequency by treating the extracted phase of the wavelet transform as shown in equation \@ref(eq:rho). Thereby I detect when phases were aligned over different trials. This can be done for all frequencies of the wavelet transform and is an effective way of quantifying phase lock using trial based data (Figure \@ref(fig:WaveletRhoNested)). The activation of the MS lead to an increase of $\rho$ during the stimulation period and dropped off rapidly after the stimulation was finished. 


(ref:WaveletRhoNestedShortCaption) Phase lock in the PaS and MEC
(ref:WaveletRhoNestedCaption) **Phase lock in the PaS and MEC.** The phase lock was quantified as $\rho$ where a value of 1 (yellow) indicates perfect phase alignment over repeated trials and $0$ (dark blue) none. The random $\rho$ given N trials is expected to be $1/\sqrt{N_{trial}}$ (MS stimulation: $0.2$; fibre stimulation: $0.183$). The y-axis shows the frequency (Hz) and the x-axis the time. Stimulation occurs from $5s$ to $10s$ (blue line). Major columns refer to the recording location (PaS or MEC) and minor columns to the stimulation location (MS or fibre stimulation). Each row shows a different stimulation frequency ($2Hz$, $4Hz$, $8Hz$, $16Hz$, $32Hz$).

```{r WaveletRhoNested, fig.cap='(ref:WaveletRhoNestedCaption)', fig.scap='(ref:WaveletRhoNestedShortCaption)',echo=F, fig.align='center', fig.height=5, fig.width=7, warning=FALSE}
RhoWaveletOverview <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/RhoWaveletOverview.rds")
RhoLogTable <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/RhoLogTable.rds")
HDIBetaMaxRhoLogOddsSummary <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/HDIBetaMaxRhoLogOddsSummary.rds")

RhoLogTable[,`:=`(GroupComparison=RecLoc, OutOfInterval=0),]
RhoLogTable[(`2.5%`> pi/sqrt(3)*0.1&`50%`> pi/sqrt(3)*0.1&`97.5%`> pi/sqrt(3)*0.1)|(`2.5%` < -pi/sqrt(3)*0.1&`50%`< -pi/sqrt(3)*0.1&`97.5%`< -pi/sqrt(3)*0.1),OutOfInterval:=1,]

RhoWaveletOverview
```
Indeed, I could clearly detect that phases were aligned during MS stimulation at all stimulation frequencies and was even pronounced when using the $2Hz$ stimulation. Due to the sensitivity of the $\rho$ value harmonics picked up by the wavelet transform are now more pronounced. Peaks of harmonics will appear as the multiple of the original signal and can be considered artefacts due to the fact that they most likely have no physiological relevance. The peak of the phase lock was centred around the stimulation frequency and seemed to change given the stimulation frequency. Similarly to the power, a stimulation of the fibres in the PaS or the MEC did not appear to be sufficient to achieve phase lock (Figure \@ref(fig:SpectrumRhoNested)). To quantify the strength of the phase lock in relation to pre-stimulus and-post stimulus times I divided the simulation blocks into three parts as shown with the power. Then I calculated the average $\rho$ for any given frequency and stimulation block. As discussed before the $\rho$ of a random distribution can be formalised as $\rho_{random}=\frac{1}{\sqrt{trial_{N}}}$ and therefore values which are true phase lock should be significantly larger than our $\rho_{random}$ given our number of trials. We can think of $\rho$ similarly as our connectivity which was described by the $beta$ distribution as a continuous value with same limits and similar interpretation. I calculated the log odds (equation \@ref(eq:logOddsRatio)) ratio of $\rho$ and $\rho_{random}$ to eliminate false positives which could occur due to not enough trials. Hence, this approach allowed me to account for uncertainty in the estimate and transformed our values from a range of 0 and 1 to normally distributed data with possible values of $-\infty$ to $\infty$. Now a value of $0$ means that there is no change in locking compared to $\rho_{random}$, negative values will describe the log reduction and positive values a log increase of phase lock. 

(ref:SpectrumRhoNestedShortCaption) Phase lock spectrum in the PaS and MEC
(ref:SpectrumRhoNestedCaption) **Phase lock spectrum in the PaS and MEC.** Log Odds Ratio of $\rho$ and the expected random ($\rho_{random}$) during different stimulation frequencies (rows) extracted from the wavelet transform for time blocks before the stimulation (Pre - black line), during the stimulation (Stim - blue line) and directly after (Post - grey line). The dashed lines indicate the stimulation frequency whereas the continuous lines show the mean $\rho$ Log Odds Ratio with its SEM. The log odds ratio represents a log fold change in phase lock compared to $\rho_{random}$. The two left columns show responses in the PaS after septal stimulation (column 1) and stimulation of the fibres in the PaS (column 2). Analogous the two right columns show the response in the MEC. The frequency on the x-axis is log~2~ transformed. PaS with MS stimulation: 9 session/4 animals; PaS with fibre stimulation: 9 session/4 animals; MEC with MS stimulation: 7 session/4 animals; MEC with fibre stimulation: 11 session/6 animals.

```{r SpectrumRhoNested, fig.cap='(ref:SpectrumRhoNestedCaption)', fig.scap='(ref:SpectrumRhoNestedShortCaption)', echo=F, fig.align='center', fig.height=5, fig.width=7, warning=FALSE}
RhoLogOddsFreqSpectrumPlotFacetOverview <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/RhoLogOddsFreqSpectrumPlotFacetOverview.rds")
RhoLogOddsFreqSpectrumPlotFacetOverview
```

Analogue to the *in-vitro* connectivity data the region for practical equivalence for log odds ratios is $\pi/\sqrt{3}\cdot0.1$ and therefore roughly $0.183$. Strong increases of phase lock can be seen at almost all stimulation frequencies including their harmonics (Figure \@ref(fig:SpectrumRhoNested)). The log-transform confirmed our suspicion that the fibre stimulation did not induce changes of locking. However, to validate and quantify whether there was a change I used an asymptotic regression model with hierarchical structure to account for variance due to trials, animals, and sessions but also recording locations, stimulation locations, and stimulation blocks (pre/stim/post) and type (burst) given a stimulation frequency. The estimated $\mu_{\rho_{LOR}}$ during pre and post blocks did not differ from each other and across stimulation frequencies, recording or stimulation locations (Figure \@ref(fig:RhoLogOddsModelFrequency)). The asymptotic maximum of the LOR was significantly higher during MS stimulation trials in the PaS (`r HDItxt(HDIBetaMaxRhoLogOddsSummary[3])`) and the MEC (`r HDItxt(HDIBetaMaxRhoLogOddsSummary[4])`) and was outside the ROPE completely. In contrast, stimulation of fibres lead to no detectable change (PaS: `r HDItxt(HDIBetaMaxRhoLogOddsSummary[7])` and MEC: `r HDItxt(HDIBetaMaxRhoLogOddsSummary[8])`). The pre-stimulus and post-stimulus blocks showed no change suggesting that locking was not carried through for longer than the stimulation period. The stimulation frequency showed to have an effect early on at $3Hz$ `r HDISummary(RhoLogTable[OutOfInterval==1,][1:2,])` and increased steadily until it reached an estimated $\mu_{\rho_{LOR}}$ of `r HDISummary(RhoLogTable[OutOfInterval==1&StimulationFrequency==32,])` at $32Hz$. 

(ref:RhoLogOddsModelFrequencyShortCaption) $\rho$ log odds dependend on stimulation frequency
(ref:RhoLogOddsModelFrequencyCaption) $\rho$ **log odds dependend on stimulation frequency.** Log Odds Ratio of $\rho$ indicating phase lock during different stimulation frequencies. Black represents measurements before stimulation (Pre), blue during stimulation (Stim), and grey after stimulation (Post). The left half shows responses in the PaS and the right half in the MEC. The top row were measurements from MS stimulation and the bottom row from fibre stimulation in the PaS or MEC. Single dots are values from sessions. The line is the estimated mean population response with the 95\% credible interval indicated as shaded area. The frequency on the x-axis is log~2~ transformed. PaS with MS stimulation: 9 session/4 animals; PaS with fibre stimulation: 9 session/4 animals; MEC with MS stimulation: 7 session/4 animals; MEC with fibre stimulation: 11 session/6 animals.

```{r RhoLogOddsModelFrequency, fig.cap='(ref:RhoLogOddsModelFrequencyCaption)', fig.scap='(ref:RhoLogOddsModelFrequencyShortCaption)', echo=F, fig.align='center', fig.height=5, fig.width=7, warning=FALSE}
RhoLogOddsModelFrequencyPlot <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/RhoLogOddsModelFrequencyPlot.rds")
UnitSummary <- readRDS("/alzheimer/Daniel_Data/R/Thesis/Data/UnitCountSummary.rds")
RhoLogOddsModelFrequencyPlot
```

Seeing that the power and the phase was mainly driven by septal stimulation and not fibre stimulation I was wondering whether PV^+^ cells in the MS were specially linked to the PaS or whether a other PV^-^ cells were similarly set up. Due to the fact that the silicon probe in the MS was fitted with a light fibre it was possible to give short $1Hz$ $2ms$ long pules. These pulses are sufficient to drive somatic ChR2, depolarise the cell and makes them spike (Figure \@ref(fig:UnitID)A). I recorded `r UnitSummary[UnitLoc=="MS",.N,]` single units in the MS $4$ of which had to be excluded from the characterisation since they did not sufficiently spike during the $1Hz$ stimulation ($<1000$). Based on $1ms$ time bins a Z-score indicated the standardised difference in spiking and a $10ms$ after stimulation onset window was used to evaluate the Z-Score (Figure \@ref(fig:UnitID)B). Only cells which had a positive change of at least $8\sigma$ were classified as PV^+^ cells. The high value was picked to account for low firing rates and sporadic spikes during the short $1s$ window of stimulation. By using these criteria I was able to identify $9$ PV^+^ cells in the MS. 

(ref:UnitIDShortCaption) Identification of PV^+^ MS cells
(ref:UnitIDCaption) **Identification of PV^+^ MS cells.** (A) Single units were stimulated in the MS using a 1Hz 2ms long pulse of blue light (blue line). Spiking (short lines) was recorded over at least 300 trials so that regularly triggered spiking during the stimulation could be observed. (B) From spiking during the 1Hz stimulation a Z-score could be derived from the spike count using 1ms binning. Dark blue indicates a strong positive increase in firing and red a decrease. The Z-score is in $\sigma$ from the mean over the whole 1s time frame. The dashed vertical line shows the point of stimulation and the horizontal dashed line the units which had more than a $8\sigma$ increase during a 10ms window. 41 Units from 14 session/7 animals.
```{r UnitID, fig.cap='(ref:UnitIDCaption)', fig.scap='(ref:UnitIDShortCaption)', echo=F, fig.align='center', fig.height=7, fig.width=7, warning=FALSE}
ChR2UnitResponse <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/ChR2UnitResponse.rds")
ChR2UnitResponse
```
To see if the identified cells corresponded to a homogeneous group of cells with similar functional properties I calculated the spike auto correlogram (Figure \@ref(fig:UnitDiversity)A). Indeed for the fraction of cells which were recorded with an LFP in the PaS cells showed distinguishable burst properties and firing frequencies. Bursting cells can be identified by their sharp peaks close to the refractory period represented by the gap in the spike correlogram. Also some of the cells had oscillatory firing behaviour whilst others did not (Figure \@ref(fig:UnitDiversity)B). This can be seen by regular bumps in the auto correlogram which can roughly be found at $0.2s$ intervals. When looking at the spike timing of PV cells in relation to the PaS $\theta$-phase cells did not show a clear coherent phase preference (Figure \@ref(fig:UnitDiversity)B) as a group. However, cells did fire a specific phases of the LFP with some preferentially spiking at the peak ($90^{\circ}$) and others at the trough $-90^{\circ}$ with different degrees of locking Figure \@ref(fig:UnitDiversity)C). All of this underlines that PV^+^ in the MS are a non-homologous group with distinct firing preferences and potentially functional implications.

(ref:UnitDiversityShortCaption) Diversity in PV^+^ MS single units
(ref:UnitDiversityCaption) **Diversity in PV^+^ MS single units.** (A) Spike auto correlogram for PV^+^ single units in the MS show refractory period as gap and in twow cells dominant bursting behaviour indicated by the large peaks at the centre. (B) A larger window of the same cells indicates different oscillatory behaviours. (C) The corresponding PaS $\theta$-phase preference for the units range from the peak of the $\theta$-oscillation ($90^{\circ}$) to the trough ($-90^{\circ}$).
```{r UnitDiversity, echo=F, fig.align='center', fig.cap='(ref:UnitDiversityCaption)', fig.height=6, fig.scap='(ref:UnitDiversityShortCaption)', fig.width=9, message=FALSE, warning=FALSE}
UnitDiversity <- readRDS(file = "/alzheimer/Daniel_Data/R/Thesis/Data/UnitDiversity.rds")
UnitDiversity
```
