# References {-}

<!-- <div id="refs"></div> -->
\small\begingroup
\noindent
\vspace{-2em}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.3in}
\setlength{\parskip}{2.5pt}

<div id="refs"></div>

\endgroup

<!-- \clearpage -->

\backmatter
# Appendix {- #appendix}
## Additional tables {-}
```{r CellParTable,tidy=FALSE, echo=FALSE, warning=F}
ParameterTable <- readRDS(file = "Data/ParameterTable.rds")
#ParameterTable[,Parameter:=gsub(x = Parameter, pattern = "Ω", "$\\\\Omega$"),]#ParameterTable[Parameter%in%c("AHP (mV)", "AP-Height (mV)", "Cm (pF)", "Firing Frequency (Hz)", "Rheobase (pA)", "Rin (MOhm)", "Threshold (mV)", "Vm (mV)"), `:=`(`2.5%`=round(`2.5%`, digits = 1), `50%`=round(`50%`, digits = 1), `97.5%`=round(`97.5%`, digits = 1)),]
setorder(ParameterTable, "Parameter")
kableExtra::kbl(
  x = ParameterTable[,.(Parameter, `Cell Type`, `2.5%`, `50%`, `97.5%`),],
#  align = "c",
  caption = 'Cell Parameters',
  digits = 3, escape = T,
  longtable = TRUE,
  booktabs = TRUE) %>%
add_header_above(c(" " = 2, "HDI" = 3)) %>%
 kable_styling(font_size = 8, latex_options = c("repeat_header"), repeat_header_method = "replace", repeat_header_text = "Cell Parameters") %>%
  collapse_rows(columns = 1, latex_hline = "major", longtable_clean_cut = T)

```

```{r CellParCompTable,tidy=FALSE, echo=FALSE, warning=F}

# # knitr::kable(
# #   x = ParameterCompTable[,.(Parameter, `Comparison`, `2.5%`, `50%`, `97.5%`, `non-zero`),],
# #   caption = 'Parameters Differences',
# #   digits = 2,
# #   longtable = TRUE,
# #   booktabs = TRUE) %>% 
# # add_header_above(c(" " = 2, "HDI" = 3, "  " = 1)) %>%
# # kable_styling(latex_options = c("repeat_header"))
# 
kableExtra::kbl(
  x = ParameterCompTable[,.(Parameter, `Comparison`, `2.5%`, `50%`, `97.5%`, `non-zero`),],
  caption = 'Cell Parameter Differences',
  digits = 3,
  longtable = TRUE,
  booktabs = TRUE) %>%
add_header_above(c(" " = 2, "HDI" = 3, "  " = 1)) %>%
 kable_styling(font_size = 8, latex_options = c("repeat_header"), repeat_header_method = "replace", repeat_header_text = "Cell Parameter Differences") %>%
  collapse_rows(columns = 1, latex_hline = "major", longtable_clean_cut = T)
```

\pagebreak
## Abbreviations {-}


```{r abbreviationTable, tidy=FALSE, echo=FALSE, warning=F}
AbbTable <- data.table(Abbrevation=c("ACSF", "sACSF", "PV", "ChAT", "TTX", "4-AP", "PaS", "PrS", "MEC", "MS", "IPSC", "IPSP", "EPSC", "EPSP", "PBS", "PFA", "VR", "LFP", "NBQX", "DH$\\beta$E","MLA", "NMDA", "AMPA", "ACh", "GABA", "ChR2", "STDP", "CCK", "HCN", "LRN", "DG", "CB", "LEC", "AP", "AHP", "V\\textsubscript{m}", "C\\textsubscript{m}", "FWHM", "ISI", "FIR", "HDI", "VGAT", "FS", "DS", "IN", "WFS1", "R\\textsubscript{in}", "YFP", "ROPE", "LOR", "SAP"),
                       Meaning=c("Artificial Cerebral Spinal Fluid", "Sucrose Artificial Cerebral Spinal Fluid", "Parvalbumin", "Choline Acetyltransferase", "Tetrodotoxin", "4-Aminopyridine", "Parasubiculum", "Presubiculum", "Medial Entorhinal Cortex", "Medial Septum","Inhibitory Postsynaptic Current", "Inhibitory Postsynaptic Potential", "Excitatory Postsynaptic Current", "Excitatory Postsynaptic Potential", "Phosphate-Buffered Saline", "Paraformaldehyde", "Virtual Reality", "Local Field Potential", "2,3-Dioxo-6-nitro-1,2,3,4-tetrahydrobenzo[f]quinoxaline-7-sulfonamide", "Dihydro-$\\beta$-erythroidine hydrobromide", "Methyllycaconitine citrate", "N-Methyl-D-aspartic acid", "$\\alpha$-amino-3-hydroxy-5-methyl-4-isoxazolepropionic acid", "Acetylcholine", "$\\gamma$-aminobutyric acid", "Channelrhodopsin-2", "Spike Time Dependent  Plasticity", "Cholecystokinin", "Hyperpolarisation-Activated Cyclic Nucleotide-Gated Non-Selective Cation Channel", "Low-Rhythmic Firing Neurons", "Dentage Gyrus", "Calbindin", "Lateral Entorhinal Cortex", "Action Potential", "Afterhyperpolarisation", "Resting Membrane Potential", "Membrane Capacitance", "Full Width Half Maximum", "Inter-Spike-Interval", "Finite Impulse Response", "Highest Density Interval", "Vesicular Inhibitory Amino Acid Transporter", "Fast Spiking", "Delayed Spiking", "Interneuron", "Wolframin ER Transmembrane Glycoprotein", "Input Resistance", "Yellow Fluorescent Protein", "Region Of Practical Equivalence", "Log Odds Ratio",  "IgG192-Saporin"))
### table over more than a page --> warning should be fixed
setorder(x = AbbTable, "Abbrevation")
knitr::kable(
  x = AbbTable,
  caption = 'List of Abbreviations',
  col.names = NULL,
  label = NA,
  longtable = TRUE,
  booktabs = TRUE,
  format = "simple",
  escape = F)
```

\listoffigures
\pagebreak
## Statement of contributions {-}
All experiments in this thesis were conceived and designed by my PhD supervisor Prof. Dr. Dietmar Schmitz and myself. All surgeries, *in-vivo* recordings and data analysis were performed by me.
\vspace{0.5cm}
\hspace{0pt}

\noindent
Detailed contributions of other people to the data presented in this work include:

\vspace{1cm}
\hspace{0pt}

\noindent
**Dr. Rosanna Sammons** (AG Schmitz, Neuroscience Research Centre, Charité Berlin) contributed 81/214 recordings of parasubicular cells.

\noindent
**Dr. Alexander Stumpf** (AG Schmitz, Neuroscience Research Centre, Charité Berlin) contributed 29/214 recordings of parasubicular cells.

\noindent
**Susanne Rieckmann** and **Anke Schönherr** organised breeding and performed genotyping of transgenic animals.

\pagebreak
## List of publications {-}

\hypertarget{Relatedtothisdissertation}{%
\subsection*{Related to this dissertation}\label{Relatedtothisdissertation}}
\addcontentsline{toc}{subsection}{Related to this dissertation}

\small\begingroup{

\noindent
\vspace{-2em}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.3in}
\setlength{\parskip}{2.5pt}

\hypertarget{refs}{}

\leavevmode\hypertarget{}{}
Sammons RP, \textbf{Parthier D}, Stumpf A, Schmitz D (2019) Electrophysiological and Molecular Characterization of the Parasubiculum. The Journal of Neuroscience: The Official Journal of the Society for Neuroscience 39:8860--8876.

\leavevmode\hypertarget{}{}
Gonzalez-Sulser A, \textbf{Parthier D}, Candela A, McClure C, Pastoll H, Garden D, Sürmeli G, Nolan MF (2014) GABAergic projections from the medial septum selectively inhibit interneurons in the medial entorhinal cortex. J Neurosci 34:16739--16743.

}
\endgroup

<!-- \vspace{1cm} -->

\hypertarget{Non-relatedtothisdissertation}{%
\subsection*{Non-related to this dissertation}\label{Non-relatedtothisdissertation}}
\addcontentsline{toc}{subsection}{Non-related to this dissertation}

\small\begingroup{

\noindent
\vspace{-2em}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.3in}
\setlength{\parskip}{2.5pt}

\hypertarget{refs}{}
\leavevmode\hypertarget{}{}
Beed P, Ray S, Velasquez LM, Stumpf A, \textbf{Parthier D}, Swaminathan A, Nitzan N, Breustedt J, Las L, Brecht M, Schmitz D (2020) Species-specific differences in synaptic transmission and plasticity. Scientific Reports 10:16557.

\leavevmode\hypertarget{}{}
Kornau H-C, Kreye J, Stumpf A, Fukata Y, \textbf{Parthier D}, Sammons RP, Imbrosci B, Kurpjuweit S, Kowski AB, Fukata M, Prüss H, Schmitz D (2020) Human Cerebrospinal Fluid Monoclonal LGI1 Autoantibodies Increase Neuronal Excitability. Annals of Neurology 87:405--418.

\leavevmode\hypertarget{}{}
Stumpf A, \textbf{Parthier D}, Sammons RP, Stempel AV, Breustedt J, Rost BR, Schmitz D (2018) Cannabinoid type 2 receptors mediate a cell type-specific self-inhibition in cortical neurons. Neuropharmacology 139:217--225.

\leavevmode\hypertarget{}{}
\textbf{Parthier D}, Kuner T, Körber C (2018) The presynaptic scaffolding protein Piccolo organizes the readily releasable pool at the calyx of Held. The Journal of Physiology 596:1485--1499.

\leavevmode\hypertarget{}{}
Rueckl M, Lenzi SC, Moreno-Velasquez L, \textbf{Parthier D}, Schmitz D, Ruediger S, Johenning FW (2017) SamuROI, a Python-Based Software Tool for Visualization and Analysis of Dynamic Time Series Imaging at Multiple Spatial Scales. Frontiers in Neuroinformatics 11:44.

\leavevmode\hypertarget{}{}
Greifzu F, \textbf{Parthier D}, Goetze B, Schlüter OM, Löwel S (2016) Ocular Dominance Plasticity after Stroke Was Preserved in PSD-95 Knockout Mice. PLOS ONE 11:e0149771.

\leavevmode\hypertarget{}{}
Zhang W, Schmelzeisen S, \textbf{Parthier D}, Frings S, Möhrlen F (2015) Anoctamin Calcium-Activated Chloride Channels May Modulate Inhibitory Transmission in the Cerebellar Cortex. PLOS ONE 10:e0142160.

}
\endgroup
\newpage
\pagebreak


## Eidestattliche Erklärung {-}
\vspace*{2in}
Hiermit versichere ich,
\vspace{11pt}

\noindent
Daniel Parthier (Matrikelnummer: 5144911) geboren am 05.08.1987 in Potsdam,
an Eides statt, dass die vorliegende Arbeit von mir selbstständig und ohne unerlaubte Hilfe Dritter verfasst wurde und ich keine anderen als die angegebenen Quellen und Hilfsmittel verwendet sowie wörtliche und sinngemäße Zitate als solche kenntlich gemacht habe.

\noindent
Diese Arbeit hat in gleicher oder ähnlicher Form noch keiner anderen Prüfungsbehörde vorgelegen und wurde bisher nicht veröffentlicht.
\vspace{24pt}

\noindent
Berlin, 28.09.2021
\vspace*{\fill}